---
title: "Reconnaissance tracking"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
  fig.path = "figures/",
  out.width = "100%"
)
```

```{r setup, include=FALSE,echo=FALSE}
library(bohemia)
library(tidyverse)
library(yaml)
library(databrew)
library(leaflet)
library(kableExtra)
library(ggplot2)
theme_set(theme_simple())
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# read in credentials
creds <- read_yaml('../credentials/credentials.yaml')
form_name_mz <- 'recon'
form_name_tz <- 'recon_geo'

data_file <- 'data.RData'

if(data_file %in% dir()){
  load(data_file)
} else {
  # read in moz data
  recon_mz <- odk_get_data(
    url = creds$moz_odk_server,
    id = form_name_mz,
    id2 = NULL,
    unknown_id2 = FALSE,
    uuids = NULL,
    exclude_uuids = NULL,
    user = creds$moz_odk_user,
    password = creds$moz_odk_pass
  )
  
  # get non repeat data
  recon_mz <- recon_mz[[2]]
  
  # read in tz data
  recon_tz <- odk_get_data(
    url = creds$tza_odk_server,
    id = 'recon',
    id2 = NULL,
    unknown_id2 = FALSE,
    uuids = NULL,
    exclude_uuids = NULL,
    user = creds$tza_odk_user,
    password = creds$tza_odk_pass
  )
  
  # get non repeat data
  recon_tz <- recon_tz[[2]]
  
  save(recon_tz,
       recon_mz,
       file = data_file)
}

# change device id to numeric
recon_tz$device_id <- as.character(recon_tz$device_id)

# join tz and mz data 
recon_data <- bind_rows(recon_tz, recon_mz)

# rm(recon_tz, recon_mz)

# get data data 
recon_data$date <- as.Date(strftime(recon_data$start_time, format = "%Y-%m-%d"))

# extract lat long
recon_data$lon <- as.numeric(unlist(lapply(strsplit(recon_data$location, ' '), function(x) x[1])))
recon_data$lat <- as.numeric(unlist(lapply(strsplit(recon_data$location, ' '), function(x) x[2])))

# get indicator for if location has been geocoded
recon_data$geo_coded <- ifelse(!is.na(recon_data$lon) | !is.na(recon_data$lat), TRUE, FALSE)

```


# Total number of submitted forms by geographic level

### By country

```{r}
recon_data %>%
  group_by(Country) %>%
  summarise(Total = n()) %>% databrew::prettify()
```

### By district

```{r}
recon_data %>%
  group_by(District) %>%
  summarise(Total = n()) %>% databrew::prettify()
```


### By village

```{r}
recon_data %>%
  group_by(Village) %>%
  summarise(Total = n()) %>% databrew::prettify()
```

### By hamlet

```{r}
recon_data %>%
  group_by(Country, District, Village, Hamlet) %>%
  summarise(Total = n()) %>% databrew::prettify()
```

# Pending forms

The below table shows the total number of hamlets by country:

```{r}
right <- locations %>% group_by(Country) %>% summarise(Total = n())
databrew::prettify(right)
```

### Pending forms by Country

```{r}
recon_data %>% group_by(Country) %>%
  summarise(Done = n()) %>% left_join(right) %>%
  mutate(`Percent finished` = round(Done / Total * 100, digits = 2)) %>% databrew::prettify()
```

### Pending forms by District

```{r}
right <- locations %>% group_by(Country, District) %>% summarise(Total = n())
recon_data %>% group_by(Country, District) %>%
  summarise(Done = n()) %>% left_join(right) %>%
  mutate(`Percent finished` = round(Done / Total * 100, digits = 2)) %>% databrew::prettify()
```

### Pending forms by Village

```{r}
right <- locations %>% group_by(Country, District, Village) %>% summarise(Total = n())
recon_data %>% group_by(Country, District, Village) %>%
  summarise(Done = n()) %>% left_join(right) %>%
  mutate(`Percent finished` = round(Done / Total * 100, digits = 2)) %>% databrew::prettify()
```

# List of all hamlets with status

```{r}
right <- locations %>% group_by(Country, District, Village, Hamlet) %>% summarise(Total = n())
recon_data %>% group_by(Country, District, Village) %>%
  summarise(Done = n()) %>% left_join(right) %>%
  mutate(`Status` = ifelse(Done >= Total, 'Done', 'Not done')) %>% 
  dplyr::select(-Done, -Total) %>%
  dplyr::select(Status, Hamlet, Village, District, Country) %>%
  databrew::prettify()
```

# Number of forms by FW

```{r}
file_name <- 'fids.RData'
if(file_name %in% dir()){
  load(file_name)
} else {
  registered_workers_tza <- gsheet::gsheet2tbl("https://docs.google.com/spreadsheets/d/1o1DGtCUrlBZcu-iLW-reWuB3PC8poEFGYxHfIZXNk1Q/edit#gid=0")
    registered_workers_moz <- gsheet::gsheet2tbl("https://docs.google.com/spreadsheets/d/1o1DGtCUrlBZcu-iLW-reWuB3PC8poEFGYxHfIZXNk1Q/edit#gid=490144130")
    registered_workers_other <- gsheet::gsheet2tbl("https://docs.google.com/spreadsheets/d/1o1DGtCUrlBZcu-iLW-reWuB3PC8poEFGYxHfIZXNk1Q/edit#gid=179257508")
    fids <- bind_rows(registered_workers_tza,
                      registered_workers_moz %>% dplyr::select(-tablet_id),
                      registered_workers_other)
    save(fids, file = file_name)
}

pd <- recon_data %>%
  group_by(bohemia_id = wid) %>%
  summarise(`Forms filled` = n()) %>%
  ungroup %>%
  left_join(fids) %>% dplyr::select(bohemia_id, First = first_name, Last = last_name,
                                    `Forms filled`) %>%
  mutate(Location = ifelse(bohemia_id <= 300, 'TZA', ifelse(bohemia_id <= 600, 'MOZ', 'Not-identified')))
databrew::prettify(pd)

```


# Total forms submitted over time

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# get submissions over time and plot
plot_data <- recon_data %>% group_by(Country,date) %>% summarise(counts = n()) 
plot_data$cum_sum <- cumsum(plot_data$counts)

ggplot(plot_data, aes(date, cum_sum, color = Country)) + 
  geom_line() +
  labs(x = 'Date',
       y = 'Cumulative recon forms submitted',
       title = 'Cumulative forms over time') +
  theme_databrew()
```

# Total forms submitted by geography

### By country

```{r}
# get total submissions
ggplot(recon_data %>% group_by(Country) %>% summarise(counts = n()), aes(Country, counts)) +
  geom_bar(stat = 'identity', alpha = 0.7) +
  labs(x = '', 
       y = '# forms submitted',
       title = 'Total forms submitted') + 
  theme_databrew()
  
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
plot_data <- recon_data %>% group_by(Country,Region, District, Ward, Village, Hamlet) %>% 
  summarise(counts = n())
```

### By region

```{r}
ggplot(plot_data, aes(Region, counts, fill = Country)) + geom_bar(stat = 'identity',alpha = 0.7) +
  labs(x = '', 
       y = '# forms submitted',
       title = 'Total forms submitted') + 
  theme_databrew()
```


### By district

```{r}
ggplot(plot_data, aes(District, counts, fill = Country)) + geom_bar(stat = 'identity',alpha = 0.7) +
  labs(x = '', 
       y = '# forms submitted',
       title = 'Total forms submitted') + 
  theme_databrew()
```

### By ward

```{r}

ggplot(plot_data, aes(Ward, counts, fill = Country)) + geom_bar(stat = 'identity',alpha = 0.7) +
  labs(x = '', 
       y = '# forms submitted',
       title = 'Total forms submitted') + 
  theme_databrew() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.7)) 
```


# Percent of forms geocoded in each country

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# percent of locations with geo code
plot_data <- recon_data %>% group_by(Country)  %>% summarise(counts = n(),
                                                            num_geo_coded = sum(geo_coded == TRUE),
                                                            percent_geo_coded = round((num_geo_coded/counts)*100, 2))

ggplot(plot_data, aes(Country, percent_geo_coded)) + geom_bar(stat = 'identity',alpha = 0.7) +
  labs(x = '', 
       y = '% of forms',
       title = 'Percent of forms submitted with geocoding') + 
  theme_databrew() 

```

# Map of locations that were geocoded

```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # BECAUSE HERE THERE IS A LAT OF -122 IN PALO ALTO.
 m <- leaflet() %>%
        addTiles()  
 m <- addCircleMarkers(map = m, 
                              lng=recon_data$lat, 
                              lat=recon_data$lon,
                              radius = 4, 
                              opacity = 1) 
 
m
```


# Table of all submitted forms for both countries

```{r, echo=FALSE, warning=FALSE, message=FALSE}
kableExtra::kable(recon_data,format = 'html', align = "lccrr") %>%
    kable_styling() %>%
    scroll_box(width = "1000px", height = "500px")

```