---
title: "Asociation of GPS watch ownership and spatial awareness"
params:
  df: data.frame()
date: ''
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble.sty
    keep_tex: yes
    latex_engine: xelatex
  word_document: default
fig_width: 5
fig_height: 3.5
bibliography: bibliography.bib
---


```{r setup, include=FALSE}
# output: 
#   pdf_document:
#     latex_engine: xelatex
#     template: pdf/layout.tex
#     includes:
#       in_header: pdf/preamble.sty
library(knitr)# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = F,
               fig.height = 3.5,
               fig.width = 5)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
library(tidyverse)
library(extrafont)
library(Rmisc)
extrafont::loadfonts()
source('functions.R')
# df <- params$df
load('df.RData')

make_authors <- function(

  authors = data.frame(name = c('Joe Brew',
                                'Eldo Elobolobo'),
                       email = c('rainbowunicorn@aol.com',
                                 'fairypixels@hotmail.com'),
                       affilitation = c("isglobal,vu",
                                        "isglobal,cism"),
                       footnote = c("Corresponding Author",
                                    rep('', 1)),
                       stringsAsFactors = FALSE),

  institutions = data.frame(code = c('isglobal',
                                     'icl',
                                     'cism',
                                     'vu',
                                     'uva',
                                     'ma'),
                            institution = c('Barcelona Institute for Global Health',
                                            'Imperial College London',
                                            'Centro de Investigação em Saúde de Manhiça',
                                            'VU University Amsterdam',
                                            'University of Amsterdam',
                                            'Maragra Açucar SA, Subsidiary of Illovo Sugar Ltd'),
                            country = c('Spain', 'UK', 'Mozambique', 'Netherlands', 'Netherlands', 'Mozambique'),
                            address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
                                        'South Kensington Campus, London SW7 2AZ, U.K.',
                                        'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
                                        'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
                                        'REC E, Roetersstraat 11, Amsterdam, Netherlands',
                                        'CP 2789, Maputo, Mozambique'),
                            stringsAsFactors = FALSE),
  cat_it = TRUE,
  include_address = TRUE,
  include_country = TRUE,
  seperator = '\n'){

  out <- c()
  for (i in 1:nrow(authors)){
    this_author <- authors$name[i]
    these_institutions <- authors$affilitation[i]
    these_institutions <- unlist(strsplit(these_institutions, ','))
    author_text <- paste0(this_author)
    for (j in 1:length(these_institutions)){
      this_institution <- these_institutions[j]
      if(any(grepl(this_institution, out))){
        author_text <- paste0(author_text,
                              '\\footrecall{',
                              this_institution,
                              '}'
        )
      } else {
        # new institution, get full name
        full_name <- institutions$institution[institutions$code == this_institution]
        address <- institutions$address[institutions$code == this_institution]
        country <- institutions$country[institutions$code == this_institution]
        author_text <- paste0(author_text,
                              '\\footremember{',
                              this_institution,
                              '}{',
                              full_name,
                              ifelse(include_address, paste0(': ', address), ''),
                              ifelse(include_country, paste0(', ', country), ''),
                              '}')
      }
    }
    author_text <- paste0(author_text, 
                          ifelse(nchar(authors$footnote[i]) > 0,
                                 paste0(' \\footnote{',
                                        authors$footnote[i],
                                        '}'),
                                 ''))
    out[i] <- author_text
  }
  out <- paste0(out, collapse = seperator)
  # cat(out)
  if(cat_it){
    cat(out)
  } else {
    return(out)
  }
}

```


\begin{center}
\begin{large}

```{r, results = 'asis', eval = TRUE}
make_authors()
```

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Abstract}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm} 

This paper provides new empirical evidence regarding the association between ownership of spatial locationing devices (wrist watches equipped with Global Positioning System, ie GPS) and accuracy in the perception of distance. Using groundbreakingly novel techniques, we show that...

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Research Highlights}
        \begin{itemize}
          \item This paper analyzes individual-level data on GPS watch ownership and spatial awareness
          \item We quantify the association through statistical significance testing
          \item This is probably the most significant scientific paper of the last decade or perhaps ever
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Keywords}
\textbf{GPS; spatial; Trump; \#yolo;}

\vspace{3mm}

\newpage

# Introduction

Does tech make people stupid or smart? Ah, the age-old question. Though conventional wisdom holds that the use of technology accelerates cognitive development [@Arellano2009], recent studies suggest that use of certain technologies may in fact result in people turning into boring vegetables [@Fink2015]. Though a great deal of literature exists on the cognitive effects of gameboys [@Lopez_Bernal_2016], shoes with flashy lights on them [@Bell2014, @Liu2010], and those nerdy hand-held miniature fans frequently used by Austrian tourists on Barcelona beaches [@Bennett_2017], no study to date has looked at the effect of GPS watches on humans' ability to correctly percetive distance [@Shretta_2017].

# Methods

We cleverly collected data from `r nrow(df)` \st{unsuspecting victims} study participants. This is not unlike the method employed by Hien [@Hien2017], but also unlikely to be not not unlike it. We basically did what Lee did [@Lee2017], but better. We are better than them. Clearly.

We asked \st{unsuspecting victims} participants to estimate the distance from Barcelona (pronounced "bar-ttthhuh-low-nuh") to two places in Africa: Rufiji, Tanzania and Mopeia, Mozambique. We used kilometers instead of miles so that people think we are smart. We than analysed (look, we used the British spelling of the word "analyse" to seem smart) the data using fancy-shmancy software [@R]. We calculated the average error (as a percentage of the correct) distance for both GPS watch owners and \st{luddites} those who don't have GPS watches, and quantified significance on the difference through a two-sample T-test (see Figure 1).

\begin{figure}[h]
\caption{Mr. T, inventor of the T-Test}
\centering
\includegraphics[width=0.2\textwidth]{t}
\end{figure}



# Results


```{r}
# Calculate average error as a percentage of true distance
df$error_rufiji <- ((df$km_rufiji - 6665) / 6665) * 100
df$error_mopeia <- ((df$km_mopeia - 7433) / 7433) * 100
df$error_rufiji_absolute <- abs(df$error_rufiji)
df$error_mopeia_absolute <- abs(df$error_mopeia)
df$avg_error_absolute <- (abs(df$error_rufiji) + abs(df$error_mopeia)) / 2
```

\st{Unsuspecting victims} Study participants' average perception of the distance from Barcelona to the two sites in question was off by approximately `r round(mean(df$avg_error_absolute, na.rm = TRUE))` percent (`r round(mean(df$error_rufiji_absolute, na.rm = TRUE))`% and `r round(mean(df$error_mopeia_absolute, na.rm = TRUE))`% for Rufiji and Mopeia, respectively). 


```{r, fig.cap = 'Participant-perceived distance from Barcelona to sites', fig.width = 6.5}
g1 <- plot_variable(variable = 'km_rufiji', db = FALSE)
g2 <- plot_variable(variable = 'km_mopeia', db = FALSE)
Rmisc::multiplot(g1, g2, cols = 2)
```



```{r}
with_gps <- df$avg_error_absolute[df$gps == 'Has a\nGPS watch']
without_gps <- df$avg_error_absolute[df$gps == 'Does not have\na GPS watch']
tt <- t.test(x = with_gps,
             y = without_gps)
is_significant <- tt$p.value < 0.05
significance_text <- ifelse(is_significant,
                           'reaches the level of statistical significance',
                           'does not reach the level of statistical significance')
```

\textbf{Differences in spatial perception by GPS watch ownership:} On average, the error in spatial distance estimation among GPS watch-owners was `r round(mean(with_gps, na.rm = TRUE), digits = 2)`%, whereas among non-GOPS watch-owners it was `r round(mean(without_gps, na.rm = TRUE), digits = 2)`%.

With a p-value of `r round(tt$p.value, digits = 4)`, the difference in error in spatial perception between owners of GPS watches and those who do not own GPS watches `r significance_text`.  The below chart shows the distribution of responses

```{r, fig.cap = 'Distribution of average error in perceived distance by GPS watch ownership'}
# plot_variable_by(variable = 'avg_error_absolute',variable_by = 'gps',
#                  db = FALSE) +
#   theme(legend.position = 'bottom',
#         legend.justification = 'center')

ggplot(data = df,
       aes(x = gps,
           y = avg_error_absolute)) +
  geom_boxplot(fill = 'darkblue', alpha = 0.2) +
  geom_point() +
  labs(x = '',
       y = 'Average error (%)')
```

# Conclusion

This paper was so good that the results speak for themselves. No conclusion necessary. Mic drop.


# Bibliography

