---
title: 'Livestock ownership and malaria risk in Mopeia'
output: github_document
---

```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8.64,
               fig.height = 4.86,
               fig.path = 'figures/')
```


```{r}
# Libraries
library(scales)
library(databrew)
library(ggplot2)
library(cism)
library(readxl)
library(tidyr)
library(tidyverse)
```

```{r}
## Comments from Carlos
# It would be (a) a heat map of livestock (only pigs, goats and poultry) at the district and (b) a heat map of malaria incidence in the cohort children finally (c) both combined in a single index. We could even try to get mosquito densities but lets start with just malaria + livestock.

# Read in the livestock data sent from Charfudin
livestock <- read_excel('data/Distribuicao de gado em Mopeia (1).xlsx', skip = 2)

# Forward fill the Posto Administrativo variable
livestock <- livestock %>%
  fill(`Posto Administrativo`, .direction = 'down')
# Remove the unecessary "Total" row
livestock <- livestock %>% dplyr::filter(`Posto Administrativo` != 'Total')
livestock <- livestock %>% gather(key, value, Bovino:Ovino)
# Translate to English
livestock_dict <- tibble(key = c('Bovino',
                                 'Caprino',
                                 'Ovino',
                                 'Suino'),
                         animal = c('Cattle',
                                    'Goats',
                                    'Sheep',
                                    'Swine'))
livestock <- livestock %>% left_join(livestock_dict, by = 'key')

```

# Major caveat

Livestock data is only available at the aggregated "localidade" level. The methods used in the production of "heat maps" in this report may be affected by the (false) concentration of livestock at a point location (the localidade) rather than their true (dispersed) locations. In other words, interpretation of heat maps should be carried out with caution as there may exist visual artefacts from the (false) concentration of livestock at the localidade centroids.

# Analysis of aggregate data

## Overlap of livestock ownership

### Total livestock ownership (all localidades)

```{r, fig.height = 8}
pd <- livestock %>% group_by(animal) %>%
  summarise(value = sum(value)) %>%
  mutate(p = paste0(round(value / sum(value) * 100, digits = 1), '%'))

cols <- RColorBrewer::brewer.pal(n = length(unique(pd$animal)),
                                 name = 'Spectral')

ggplot(data = pd,
       aes(x = animal,
           y = value,
           fill = animal)) +
  geom_bar(stat = 'identity',
           alpha = 0.8) +
  theme_databrew() +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_text(size = 20)) +
  geom_text(aes(label = paste0(comma(value), '\n(', p, ')')),
            nudge_y = -800,
            color = 'white',
            size = 5) +
  scale_fill_manual(name = '',
                    values = cols) +
  theme(legend.position = 'none')
```

### Total livestock ownership (by "Posto Administrativo")

```{r}
pd <- livestock %>% group_by(animal,`Posto Administrativo`) %>%
  summarise(value = sum(value)) %>%
  ungroup %>%
  group_by(`Posto Administrativo`) %>%
  mutate(p = paste0(round(value / sum(value) * 100, digits = 1), '%'))

ggplot(data = pd,
       aes(x = animal,
           y = value,
           fill = animal)) +
  facet_wrap(~`Posto Administrativo`) +
  geom_bar(stat = 'identity',
           alpha = 0.8) +
  theme_databrew() +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5),
        strip.text = element_text(size = 20)) +
  geom_text(aes(label = paste0(comma(value), '\n(', p, ')')),
            nudge_y = -300,
            color = 'white',
            size = 3) +
    scale_fill_manual(name = '',
                    values = cols) +
  theme(legend.position = 'none')
```


### Total livestock ownership (by "Localidade")

```{r}
pd <- livestock %>% group_by(animal,Localidade) %>%
  summarise(value = sum(value)) %>%
  ungroup %>%
  group_by(Localidade) %>%
  mutate(p = paste0(round(value / sum(value) * 100, digits = 1), '%'))

ggplot(data = pd,
       aes(x = animal,
           y = value,
           fill = animal)) +
  facet_wrap(~Localidade,
             ncol = 4) +
  geom_bar(stat = 'identity',
           alpha = 0.8) +
  theme_databrew() +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5),
        strip.text = element_text(size = 15)) +
  geom_text(aes(label = paste0(comma(value), '\n(', p, ')')),
            nudge_y = 500,
            color = 'black',
            alpha = 0.6,
            size = 3) +
    scale_fill_manual(name = '',
                    values = cols) +
  theme(legend.position = 'none')
```

### Correlation between ownership of one animal and another

On May 27 2019, Cassidy Rist asked the following (via email):

```
I wonder what the overlap is among livestock ownership. 
For example, do most people own pigs and goats, or is it 
more likely that one or the other species is owned?
```

As of now, we only have aggregate-level data (most granular: localidade). Without individual-level data, the above can be addressed. Until then. we can examine the correlation at the localidade level (below), but in doing so we're committing the ecological fallacy.

The below aims to address the above question, using aggregated data ("localidade"-level).

#### Cattle and Goats

```{r}
pd <- livestock %>% dplyr::select(Localidade, animal, value) %>% spread(key = animal, value = value)

ggplot(data = pd,
       aes(x = Cattle,
           y = Goats,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Cattle and Goats',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)')
```



#### Cattle and Sheep

```{r}

ggplot(data = pd,
       aes(x = Cattle,
           y = Sheep,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Cattle and Sheep',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') +
  ylim(0, 610)
```

#### Cattle and Swine

```{r}

ggplot(data = pd,
       aes(x = Cattle,
           y = Swine,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Cattle and Swine',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```


#### Goats and Sheep

```{r}

ggplot(data = pd,
       aes(x = Goats,
           y = Sheep,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Goats and Sheep',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```

#### Goats and Swine

```{r}

ggplot(data = pd,
       aes(x = Goats,
           y = Swine,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Goats and Swine',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```

#### Sheep and Swine

```{r}

ggplot(data = pd,
       aes(x = Sheep,
           y = Swine,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Sheep and Swine',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```


# Analysis of person-level data




```{r}
library(cism)

# Create acd from 3 datasets, per Eldo's instructions
acda <- read_csv('data/from_eldo/COST_ACD_Core_13-05-2019.csv')
acdb <- read_csv('data/from_eldo/COST_ACD_Childs_13-05-2019.csv')
acdc <- read_csv('data/from_eldo/COST_ACD_nets_13-05-2019.csv')

acd <- 
  left_join(acdb,
            acda,
            by = c('_PARENT_AURI' = '_URI')) %>%
  left_join(acdc,
            by = c('_URI' = '_PARENT_AURI'))

# Read in eldo files
eldo_perm <- read_csv('data/from_eldo/COST_Permids.New&Old_EE.csv')
eldo_census_agg <- read_csv('data/from_eldo/COST_SprayStatus_by_Village_Id_11.04.2019.EE.csv')
eldo_livestock <- read_csv('data/from_eldo/COST_ACD_Core_13-05-2019.csv')
eldo_census_2016 <- read_csv('data/from_eldo/Census_2016.csv')
eldo_census_2017 <- read_csv('data/from_eldo/COST_Censo2017_Core.1.4.2019.csv')
# census <- eldo_census_2016 %>% dplyr::select(gpc_lng,
#                                              gpc_lat,
#                                              locality_Final,
#                                              `village number_final`,
#                                              household_number) %>%
#   dplyr::rename(lng = gpc_lng,
#            lat = gpc_lat,
#            localidade = locality_Final,
#            hhid = household_number,
#            village_number = `village number_final`) %>% mutate(year = 2016) %>%
#   mutate(family_id )
#   bind_rows(eldo_census_2017 %>%
#               dplyr::rename(lng = LOCALITY_GPS_LNG,
#                             lat = LOCALITY_GPS_LAT,
#                             localidade = LOCAL_VILLAGENAME,
#                             village_number = FAMILY_ID) %>%
#               dplyr::select(village_number,
#                             localidade, lng, lat) %>%
#               mutate(village_number = as.numeric(unlist(lapply(strsplit(village_number, '-'), function(x){x[1]})))) %>%
#               mutate(year = 2017))

# Sort out locations of clusters, etc.
locations <- eldo_census_2016 %>%
  group_by(lng = gpc_lng,
           lat = gpc_lat,
           localidade = locality_Final,
           village_number = `village number_final`) %>%
  tally %>%
  dplyr::filter(n == 1) %>%
  dplyr::select(village_number, localidade, lng, lat)

# Try to get locations for each localidade for livestock
avg_locations <- locations %>%
  # mutate()
  group_by(localidade) %>%
  summarise(lng = mean(lng),
            lat = mean(lat))


# # Define function for converting a meta_instance_name from livestock
# to a household id in the census dataset
numeric_convert <- function(x){
  # Minimize
  x <- strsplit(x, '-')
  # Make numeric
  x <- lapply(x, function(z){as.character(as.numeric(z))})
  x <- lapply(x, function(z){paste0(z[1], '-', z[2], '-', z[3], '-', z[4])})
  x <- unlist(x)
  return(x)
}

# Join acd and eldo perm info
acd <-
  left_join(acd %>% 
              mutate(permid = numeric_convert(PERM_ID)), 
            eldo_perm %>% 
              mutate(permid = numeric_convert(new_permid)),
            by = 'permid')

# It would be (a) a heat map of livestock (only pigs, goats and poultry) at the district and (b) a heat map of malaria incidence in the cohort children finally (c) both combined in a single index. We could even try to get mosquito densities but lets start with just malaria + livestock.

# USE ACD data for both human distribution and animal distribution
# The relevant variables are main_animal_<animal name>

# Keep only relevant vars

acd <- acd %>%
  dplyr::select(
    permid,
         gender,
         cluster,
         longitude,
         latitude,
    chicken = MAIN_ANIMAL_CHICK,
         animals = MAIN_ANIMALS,
         cattle = MAIN_ANIMAL_CATTLE,
         pigs = MAIN_ANIMAL_PIGS,
         horses = MAIN_ANIMAL_HORSE,
         sheel = MAIN_ANIMAL_SHEEP,
         goats = MAIN_ANIMAL_COATS,
         FAMILY_ID,
         DEMOGRAFIC_DOB,
         HOME_WORK,
         HOME_WORK_TIME,
         HAD_MALARIA,
         INDUSTRY,
         MAIN_MALARIA_CONTROL_UNDERNET_SLEEP) %>%
  filter(!is.na(longitude),
         !is.na(latitude))
```

## Overall distribution of humans (point map)

(The below uses census data)

```{r}
ggplot() +
  geom_polygon(data = mop2_fortified,
               aes(x = long,
                   y = lat,
                   group = group)) +
  theme_databrew() +
  coord_map() +
  geom_point(data = locations,
             aes(x = lng,
                 y = lat),
             color = 'red',
             alpha = 0.6,
             size = 0.1) +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

## Overall distribution of humans ("heat map")

(The below uses census data)

```{r}
library(raster)
library(rasterVis)

make_grid <- function(the_densities = locations,
                      the_map = mop2,
                     griddy = 0.005){
  
  # Define the densities
  df_spatial <- the_densities %>% mutate(x = lng, y = lat) %>% filter(!is.na(lng), !is.na(lat))
  coordinates(df_spatial) <- ~x+y
  proj4string(df_spatial) <- proj4string(the_map)
  # First we need to create a uniform grid
  # Create a gridded dataframe with values 
  # for the entire range (bbox) of magude
  df_grid <- expand.grid(lng = seq(bbox(the_map)[1,1],
                                   bbox(the_map)[1,2],
                                   by = griddy), # make this smaller for better quality
                         lat = seq(bbox(the_map)[2,1],
                                   bbox(the_map)[2,2],
                                   by = griddy), # make this smaller for better quality
                         pos_score = NA)
  df_grid$latitude <- df_grid$lat
  df_grid$longitude <- df_grid$lng
  coordinates(df_grid) <- ~longitude+latitude
  proj4string(df_grid) <- proj4string(the_map)
  # # Visualize the grid
  # plot(the_map)
  # points(df_grid$lng, df_grid$lat, pch = ".")
  
  
  # Loop through each point in our grid
  for (i in 1:nrow(df_grid)){
    message(i)
    # Get distance from this grid point to every point in df_spatial
    distances <- spDistsN1(pts = df_spatial,
                           pt = df_grid[i,],
                           longlat = TRUE)
    # Define which are acceptably close (define a radius)
    close_enough <- which(distances <= 500)
    # Get a positivity score ( 1 means yes and 0 means no)
    # positivity <- stats::weighted.mean(
    #   x = df_spatial$pcr_mujer[close_enough],
    #   w = (1 / distances[close_enough])^2,
    #                                    na.rm = TRUE)
    positivity <- mean(1 / (distances[close_enough]^(1/4)), na.rm = TRUE)
    # Assign irs to the dataframe
    df_grid$pos_score[i] <- positivity
  }
  
  # Convert df_grid to raster
  temp <- df_grid@data %>% arrange(lng, lat)
  r <- rasterFromXYZ(temp[, c('lng', 'lat', 'pos_score')])
  # plot(r)
  # plot(the_map, add = TRUE)
  
  # Make grid smaller and then reconvert to raster
  proj4string(df_grid) <- proj4string(the_map)
  x <- over(df_grid, polygons(the_map))
  df_grid_small <- df_grid[!is.na(x),]
  # plot(df_grid_small,
  #      col = adjustcolor('black', alpha.f = 0.2))
  temp <- df_grid_small@data %>% arrange(lng, lat)
  r <- rasterFromXYZ(temp[, c('lng', 'lat', 'pos_score')])
  return(r)
  # # plot(r, legend = FALSE)
  # # plot(the_map, add = TRUE)
  # 
  # # plot(the_map, add = T)
}

# Define another function for getting a population adjusted grid
make_grid_adjusted <- function(the_densities = locations,
                     the_map = mop2,
                      the_population = locations,
                     griddy = 0.01){
  
  # Define the densities
  df_spatial <- the_densities %>% mutate(x = lng, y = lat) %>% filter(!is.na(lng), !is.na(lat))
  coordinates(df_spatial) <- ~x+y
  proj4string(df_spatial) <- proj4string(the_map)
  
  # Define the population densities
  population_spatial <- the_population %>% mutate(x = lng, y = lat) %>% filter(!is.na(lng), !is.na(lat))
  coordinates(population_spatial) <- ~x+y
  proj4string(population_spatial) <- proj4string(the_map)
  
  # First we need to create a uniform grid
  # Create a gridded dataframe with values 
  # for the entire range (bbox) of magude
  df_grid <- expand.grid(lng = seq(bbox(the_map)[1,1],
                                   bbox(the_map)[1,2],
                                   by = griddy), # make this smaller for better quality
                         lat = seq(bbox(the_map)[2,1],
                                   bbox(the_map)[2,2],
                                   by = griddy), # make this smaller for better quality
                         pos_score = NA)
  df_grid$latitude <- df_grid$lat
  df_grid$longitude <- df_grid$lng
  coordinates(df_grid) <- ~longitude+latitude
  proj4string(df_grid) <- proj4string(the_map)
  
  
  # Loop through each point in our grid
  for (i in 1:nrow(df_grid)){
    message(i)
    # Get distance from this grid point to every point in df_spatial
    distances <- spDistsN1(pts = df_spatial,
                           pt = df_grid[i,],
                           longlat = TRUE)
    population_distances <- spDistsN1(pts = population_spatial,
                                      pt = df_grid[i,],
                                      longlat = TRUE)
    # Define which are acceptably close (define a radius)
    close_enough <- which(distances <= 500)
    # Get a positivity score ( 1 means yes and 0 means no)
    # positivity <- stats::weighted.mean(
    #   x = df_spatial$pcr_mujer[close_enough],
    #   w = (1 / distances[close_enough])^2,
    #                                    na.rm = TRUE)
    positivity <- mean(1 / (distances[close_enough]^(1/4)), na.rm = TRUE)
    population_positivity <-   mean(1 / (population_distances[close_enough]^(1/4)), na.rm = TRUE)
    # Assign irs to the dataframe
    df_grid$pos_score[i] <- positivity / population_positivity
  }
  
  # Convert df_grid to raster
  temp <- df_grid@data %>% arrange(lng, lat)
  r <- rasterFromXYZ(temp[, c('lng', 'lat', 'pos_score')])
  # plot(r)
  # plot(the_map, add = TRUE)
  
  # Make grid smaller and then reconvert to raster
  proj4string(df_grid) <- proj4string(the_map)
  x <- over(df_grid, polygons(the_map))
  df_grid_small <- df_grid[!is.na(x),]
  # plot(df_grid_small,
  #      col = adjustcolor('black', alpha.f = 0.2))
  temp <- df_grid_small@data %>% arrange(lng, lat)
  r <- rasterFromXYZ(temp[, c('lng', 'lat', 'pos_score')])
  return(r)
  # # plot(r, legend = FALSE)
  # # plot(the_map, add = TRUE)
  # 
  # # plot(the_map, add = T)
}

heat_map <- function(r){
  # levelplot(r, contour = F, par.settings = BuRdTheme()) +
  levelplot(r, contour = F, par.settings = infernoTheme()) +
  layer(sp.polygons(mop2, lwd = 3))
}

file_name <- 'grids/1.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(griddy = 0.01)
  save(r, file = file_name)
}

heat_map(r) 
# heat_map(griddy = 0.01)

# points(df_spatial,
#        col = adjustcolor("black", alpha.f = 0.2),
#        cex = 0.2)

```

## Overall distribution of livestock (point map)

The below uses municipality data. This means that it is aggregated at the localidade-level. We also have person-level livestock data, but only in the ACD records (which are not comprehensive of all people).

```{r, fig.height = 8}
# Make a long df of livestock long
livestock_with_locations <-
  livestock %>%
  mutate(localidade = ifelse(Localidade == 'Campo sede', 'POSTO CAMPO-SEDE',
                           ifelse(Localidade == 'Catale', 'CATALE',
                                  ifelse(Localidade == 'Lua lua', 'LUA LUA',
                                         ifelse(Localidade == 'Mopeia sede', 'MOPEIA-SEDE/CUA CUA',
                                                ifelse(Localidade == 'Mungane', 'MUNGANE',
                                                       ifelse(Localidade == 'Nzanza', 'Nzanza',
                                                              ifelse(Localidade == 'Rovuma', 'ROVUMA-CONHO',
                                                                     ifelse(Localidade == 'Sambalendo',
                                                                            'SAMBALENDO/CHIMUARA',
                                                                            Localidade))))))))) %>%
  left_join(locations %>%
              dplyr::select(localidade, lng, lat) %>%
              group_by(localidade) %>%
              summarise(lng = mean(lng),
                        lat = mean(lat)))

ggplot() +
  geom_polygon(data = mop2_fortified,
               aes(x = long,
                   y = lat,
                   group = group)) +
  theme_databrew() +
  coord_map() +
  geom_point(data = livestock_with_locations,
             aes(x = lng,
                 y = lat,
                 size = value,
                 color = animal),
             alpha = 0.6) +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  facet_wrap(~animal) +
  scale_color_manual(name = '',
                     values = cols,
                     guide = FALSE) +
  scale_size_continuous(name = '',
                        breaks = c(100, 500, 1000, 2000, 3000),
                        range = c(0, 15))
```

## Overall distribution of livestock ("heat map")

The below uses municipality data. This means that it is aggregated at the localidade-level. The person-level livestock data (only in the ACD records) are not comprehensive of all people.

There are two kinds of maps for each animal: (1) a straightforward heat map (showing the distribution of animals) and (2) a human population-adjusted heatmap (showing where the animal in question is more or less prevalent, relative to the distribution of the human population). In the latter map, if livestock distribution reflected perfectly human distribution, there would be no variation in color throughout the map.

```{r}
animal_densities_list <- list()
for(i in 1:nrow(livestock_with_locations)){
  # message(i)
  this_row <- livestock_with_locations[i,]
  out <- tibble(animal = this_row$animal,
                lng = this_row$lng,
                lat = this_row$lat,
                n = 1:this_row$value)
  animal_densities_list[[i]] <- out
}
animal_densities <- bind_rows(animal_densities_list)
```

### Cattle

#### Distribution of cattle

```{r}
file_name <- 'grids/2.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(the_densities = animal_densities %>% filter(animal == 'Cattle'))
  save(r, file = file_name)
}
heat_map(r)
```

#### Distribution of cattle adjusted by human population


```{r}
file_name <- 'grids/2b.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid_adjusted(the_densities = animal_densities %>% filter(animal == 'Cattle'),
                          the_population = locations)
  save(r, file = file_name)
}
heat_map(r)
```

### Goats

#### Distribution of goats

```{r}
file_name <- 'grids/3.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(the_densities = animal_densities %>% filter(animal == 'Goats'))
  save(r, file = file_name)
}
heat_map(r)
```

#### Distribution of cattle adjusted by human population


```{r}
file_name <- 'grids/db.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid_adjusted(the_densities = animal_densities %>% filter(animal == 'Goats'),
                          the_population = locations)
  save(r, file = file_name)
}
heat_map(r)
```

### Sheep

#### Distribution of sheep

```{r}
file_name <- 'grids/4.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(the_densities = animal_densities %>% filter(animal == 'Sheep'))
  save(r, file = file_name)
}
heat_map(r)
```

#### Distribution of sheep adjusted by human population


```{r}
file_name <- 'grids/4b.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid_adjusted(the_densities = animal_densities %>% filter(animal == 'Sheep'),
                          the_population = locations)
  save(r, file = file_name)
}
heat_map(r)
```

### Swine

#### Distribution of swine

```{r}
file_name <- 'grids/5.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(the_densities = animal_densities %>% filter(animal == 'Swine'))
  save(r, file = file_name)
}
heat_map(r)
```

#### Distribution of swine adjusted by human population


```{r}
file_name <- 'grids/5b.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid_adjusted(the_densities = animal_densities %>% filter(animal == 'Swine'),
                          the_population = locations)
  save(r, file = file_name)
}
heat_map(r)
```


### All livestock

#### Distribution of all livestock

```{r}
file_name <- 'grids/all.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(the_densities = animal_densities)
  save(r, file = file_name)
}
heat_map(r)
```

#### Distribution of all livestock adjusted by human population


```{r}
file_name <- 'grids/allb.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid_adjusted(the_densities = animal_densities,
                          the_population = locations)
  save(r, file = file_name)
}
heat_map(r)
```

#### Distribution of all livestock except cattle

```{r}
file_name <- 'grids/all_not_cattle.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(the_densities = animal_densities %>% filter(animal != 'Cattle'))
  save(r, file = file_name)
}
heat_map(r)
```

#### Distribution of all livestock adjusted by human population


```{r}
file_name <- 'grids/all_not_cattleb.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid_adjusted(the_densities = animal_densities%>% filter(animal != 'Cattle'),
                          the_population = locations)
  save(r, file = file_name)
}
heat_map(r)
```




## Malaria maps

### Overall distribution of malaria (all ACD cases) (point map)


```{r}
ggplot() +
  geom_polygon(data = mop2_fortified,
               aes(x = long,
                   y = lat,
                   group = group)) +
  theme_databrew() +
  coord_map() +
  geom_jitter(data = acd,
             aes(x = longitude,
                 y = latitude),
             color = 'red',
             alpha = 0.6,
             size = 0.5) +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())
```



### Overall distribution of malaria (all ACD cases) with population (point map, jittered)

(Grey: house)
(Red: malaria case)

```{r}
ggplot() +
  geom_polygon(data = mop2_fortified,
               aes(x = long,
                   y = lat,
                   group = group)) +
  theme_databrew() +
  coord_map() +
    geom_point(data = locations,
             aes(x = lng,
                 y = lat),
             color = 'grey',
             alpha = 0.3,
             size = 0.8) +
  geom_jitter(data = acd,
             aes(x = longitude,
                 y = latitude),
             color = 'red',
             alpha = 0.6,
             size = 0.4) +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

### Overall distribution of malaria (all ACD cases) ("heat map")

```{r}
file_name <- 'grids/6.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid(the_densities = acd %>% mutate(lng = longitude, lat = latitude) %>% data.frame())
  save(r, file = file_name)
}
heat_map(r)
```

### Overall distribution of malaria, adjusted by human population (incidence heat map)

```{r}
file_name <- 'grids/7.RData'
if(file.exists(file_name)){
  load(file_name)
} else {
  r <- make_grid_adjusted(
    the_densities = acd %>% mutate(lng = longitude, lat = latitude) %>% data.frame(),
    the_population = locations)
  save(r, file = file_name)
}
heat_map(r)
```



# Map of livestock and malaria incidence combined into single index

(Excludes cattle)

The below shows the product of two values: 

1. The (percentilized) malaria distribution and

2. The (percentilized) distribution of livestock

The lighter areas show areas with the highest (percentilized) values of both. In other words, light yellow means there is a lot of malaria and livestock, whereas dark purple means the opposite.

```{r}
# Malaria
load('grids/6.RData')
malaria <- r
values(malaria) <- dplyr::percent_rank(values(malaria)) #* 100
proj4string(malaria) <- proj4string(mop2)

# Livestock
load('grids/all_not_cattle.RData')
livestock <- r
values(livestock) <- dplyr::percent_rank(values(livestock))# * 100
proj4string(livestock) <- proj4string(mop2)

# Crop
malaria <- crop(malaria, livestock)
extent(malaria) <- extent(livestock)
malaria <- projectRaster(from = malaria, 
                         to = livestock#,
                         # method = 'ngb',
                         # res = res(mosq),
                         # crs = proj4string(mosq)
  )

combined <- overlay(malaria, livestock, 
                       # fun = combine_function
                       fun = prod
                       )
heat_map(r = combined)
```

One problem with the above map is that it too closely tracks population density. There is, as would be expected, more malaria where there are more people. Accordingly, we can also examine our human population-adjusted malaria distribution and combine it (again using percentilized product aggregation) with the distribution of livestock. See below.

```{r}
# Malaria
load('grids/7.RData')
malaria <- r
values(malaria) <- dplyr::percent_rank(values(malaria)) #* 100
proj4string(malaria) <- proj4string(mop2)

# Livestock
load('grids/all_not_cattle.RData')
livestock <- r
values(livestock) <- dplyr::percent_rank(values(livestock))# * 100
proj4string(livestock) <- proj4string(mop2)

# Crop
malaria <- crop(malaria, livestock)
extent(malaria) <- extent(livestock)
malaria <- projectRaster(from = malaria, 
                         to = livestock#,
                         # method = 'ngb',
                         # res = res(mosq),
                         # crs = proj4string(mosq)
  )

combined <- overlay(malaria, livestock, 
                       # fun = combine_function
                       fun = prod
                       )
heat_map(r = combined)
```



# Pending: Map of mosquito densities

# Technical details

This document was produced on `r Sys.Date()` on a `r Sys.info()['sysname']` machine (release `r Sys.info()['release']`. To reproduce, one should take the following steps:

- Clone the repository at https://github.com/databrew/bohemia

- Populate the `analyses/livestock_distribution/data` directory with the following files: `Distribuicao de gado em Mopeia (1).xlsx` (emailed to team members from Charfudin Sacoor on May 27 2019);  

- Populate the `analyses/livestock_distribution/data/from_eldo` directory with the following files (emailed from Eldo to team on June 10 2019):

```
├── Census_2016.csv
├── COST_ACD_Core_13-05-2019.csv
├── COST_ACD_Follow_Up_Core_13-05-2019.csv
├── COST_ACD_Follow_Up_V062017_Core_10-05-2019.csv
├── COST_Censo2017_Core.1.4.2019.csv
├── COST_CROSS_2017_Core_Final_11.04.2019.csv
├── COST_CROSS_2018_Core_Final_11.10.2018.csv
├── COST_Cross_2019_Core_10-06-2019.csv
├── COST_PCD_2019_Core_15-04-2019.csv
├── COST_PCD_V062017_Core_17-01-2019.csv
├── COST_SprayStatus_by_Village_Id_11.04.2019.EE.csv
├── Livestock_Mopeia.csv
└── Pcd1_core.csv
```

- Download the `COST_Permids.New&Old_EE.csv` file, sent by Eldo to team on June 11 2019, into `data/from_eldo`.

- "Render" (using `rmarkdown`) the code in `analysis/livestock_distribution/README.Rmd`

Any questions or problems should be addressed to joe@databrew.cc