---
title: 'Livestock ownership and malaria risk in Mopeia'
output: github_document
---

```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8.64,
               fig.height = 4.86,
               fig.path = 'figures/')
```


```{r}
# Libraries
library(scales)
library(databrew)
library(ggplot2)
library(cism)
library(readxl)
library(tidyr)
library(tidyverse)
```

```{r}
## Comments from Carlos
# It would be (a) a heat map of livestock (only pigs, goats and poultry) at the district and (b) a heat map of malaria incidence in the cohort children finally (c) both combined in a single index. We could even try to get mosquito densities but lets start with just malaria + livestock.

# Read in the livestock data sent from Charfudin
livestock <- read_excel('data/Distribuicao de gado em Mopeia (1).xlsx', skip = 2)

# Forward fill the Posto Administrativo variable
livestock <- livestock %>%
  fill(`Posto Administrativo`, .direction = 'down')
# Remove the unecessary "Total" row
livestock <- livestock %>% dplyr::filter(`Posto Administrativo` != 'Total')
livestock <- livestock %>% gather(key, value, Bovino:Ovino)
# Translate to English
livestock_dict <- tibble(key = c('Bovino',
                                 'Caprino',
                                 'Ovino',
                                 'Suino'),
                         animal = c('Cattle',
                                    'Goats',
                                    'Sheep',
                                    'Swine'))
livestock <- livestock %>% left_join(livestock_dict, by = 'key')

```

# Analysis of aggregate data

## Overlap of livestock ownership

### Total livestock ownership (all localidades)

```{r}
pd <- livestock %>% group_by(animal) %>%
  summarise(value = sum(value)) %>%
  mutate(p = paste0(round(value / sum(value) * 100, digits = 1), '%'))

cols <- RColorBrewer::brewer.pal(n = length(unique(pd$animal)),
                                 name = 'Spectral')

ggplot(data = pd,
       aes(x = animal,
           y = value,
           fill = animal)) +
  geom_bar(stat = 'identity',
           alpha = 0.8) +
  theme_databrew() +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_text(size = 20)) +
  geom_text(aes(label = paste0(comma(value), '\n(', p, ')')),
            nudge_y = -800,
            color = 'white',
            size = 5) +
  scale_fill_manual(name = '',
                    values = cols) +
  theme(legend.position = 'none')
```

### Total livestock ownership (by "Posto Administrativo")

```{r}
pd <- livestock %>% group_by(animal,`Posto Administrativo`) %>%
  summarise(value = sum(value)) %>%
  ungroup %>%
  group_by(`Posto Administrativo`) %>%
  mutate(p = paste0(round(value / sum(value) * 100, digits = 1), '%'))

ggplot(data = pd,
       aes(x = animal,
           y = value,
           fill = animal)) +
  facet_wrap(~`Posto Administrativo`) +
  geom_bar(stat = 'identity',
           alpha = 0.8) +
  theme_databrew() +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5),
        strip.text = element_text(size = 20)) +
  geom_text(aes(label = paste0(comma(value), '\n(', p, ')')),
            nudge_y = -300,
            color = 'white',
            size = 3) +
    scale_fill_manual(name = '',
                    values = cols) +
  theme(legend.position = 'none')
```


### Total livestock ownership (by "Localidade")

```{r}
pd <- livestock %>% group_by(animal,Localidade) %>%
  summarise(value = sum(value)) %>%
  ungroup %>%
  group_by(Localidade) %>%
  mutate(p = paste0(round(value / sum(value) * 100, digits = 1), '%'))

ggplot(data = pd,
       aes(x = animal,
           y = value,
           fill = animal)) +
  facet_wrap(~Localidade,
             ncol = 4) +
  geom_bar(stat = 'identity',
           alpha = 0.8) +
  theme_databrew() +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5),
        strip.text = element_text(size = 15)) +
  geom_text(aes(label = paste0(comma(value), '\n(', p, ')')),
            nudge_y = 500,
            color = 'black',
            alpha = 0.6,
            size = 3) +
    scale_fill_manual(name = '',
                    values = cols) +
  theme(legend.position = 'none')
```

### Correlation between ownership of one animal and another

On May 27 2019, Cassidy Rist asked the following (via email):

```
I wonder what the overlap is among livestock ownership. 
For example, do most people own pigs and goats, or is it 
more likely that one or the other species is owned?
```

As of now, we only have aggregate-level data (most granular: localidade). Without individual-level data, the above can be addressed. Until then. we can examine the correlation at the localidade level (below), but in doing so we're committing the ecological fallacy.

The below aims to address the above question, using aggregated data ("localidade"-level).

#### Cattle and Goats

```{r}
pd <- livestock %>% dplyr::select(Localidade, animal, value) %>% spread(key = animal, value = value)

ggplot(data = pd,
       aes(x = Cattle,
           y = Goats,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Cattle and Goats',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)')
```



#### Cattle and Sheep

```{r}

ggplot(data = pd,
       aes(x = Cattle,
           y = Sheep,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Cattle and Sheep',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') +
  ylim(0, 610)
```

#### Cattle and Swine

```{r}

ggplot(data = pd,
       aes(x = Cattle,
           y = Swine,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Cattle and Swine',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```


#### Goats and Sheep

```{r}

ggplot(data = pd,
       aes(x = Goats,
           y = Sheep,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Goats and Sheep',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```

#### Goats and Swine

```{r}

ggplot(data = pd,
       aes(x = Goats,
           y = Swine,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Goats and Swine',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```

#### Sheep and Swine

```{r}

ggplot(data = pd,
       aes(x = Sheep,
           y = Swine,
           color = Localidade)) +
  geom_point(size = 4, 
             alpha = 0.8) +
  scale_color_manual(name = '',
                    values = RColorBrewer::brewer.pal(n = length(unique(pd$Localidade)), 
                                                    name = 'Spectral')) +
  theme_databrew() + 
  theme(axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  geom_line(aes(group = 1),
            stat="smooth",method = "lm", formula = y ~ 0 + I(1/x) + I((x-1)/x),
              size = 1,
              linetype ="dashed",
              alpha = 0.5) +
  labs(title = 'Association of Sheep and Swine',
       subtitle = 'Smoothed line formula: y ~ 0 + I(1/x)') 
```


# Analysis of person-level data




```{r}
library(cism)

# Create acd from 3 datasets, per Eldo's instructions
acda <- read_csv('data/from_eldo/COST_ACD_Core_13-05-2019.csv')
acdb <- read_csv('data/from_eldo/COST_ACD_Childs_13-05-2019.csv')
acdc <- read_csv('data/from_eldo/COST_ACD_nets_13-05-2019.csv')

acd <- 
  left_join(acdb,
            acda,
            by = c('_PARENT_AURI' = '_URI')) %>%
  left_join(acdc,
            by = c('_URI' = '_PARENT_AURI'))

# Read in eldo files
eldo_perm <- read_csv('data/from_eldo/COST_Permids.New&Old_EE.csv')
eldo_census_agg <- read_csv('data/from_eldo/COST_SprayStatus_by_Village_Id_11.04.2019.EE.csv')
eldo_livestock <- read_csv('data/from_eldo/COST_ACD_Core_13-05-2019.csv')
eldo_census_2016 <- read_csv('data/from_eldo/Census_2016.csv')
eldo_census_2017 <- read_csv('data/from_eldo/COST_Censo2017_Core.1.4.2019.csv')
census <- eldo_census_2016 %>% dplyr::select(gpc_lng,
                                             gpc_lat,
                                             locality_Final,
                                             `village number_final`,
                                             household_number) %>%
  dplyr::rename(lng = gpc_lng,
           lat = gpc_lat,
           localidade = locality_Final,
           hhid = household_number,
           village_number = `village number_final`) %>% mutate(year = 2016) %>%
  mutate(family_id )
  bind_rows(eldo_census_2017 %>%
              dplyr::rename(lng = LOCALITY_GPS_LNG,
                            lat = LOCALITY_GPS_LAT,
                            localidade = LOCAL_VILLAGENAME,
                            village_number = FAMILY_ID) %>%
              dplyr::select(village_number,
                            localidade, lng, lat) %>%
              mutate(village_number = as.numeric(unlist(lapply(strsplit(village_number, '-'), function(x){x[1]})))) %>%
              mutate(year = 2017))

# Sort out locations of clusters, etc.
locations <- eldo_census_2016 %>%
  group_by(lng = gpc_lng,
           lat = gpc_lat,
           localidade = locality_Final,
           village_number = `village number_final`) %>%
  tally %>%
  dplyr::filter(n == 1) %>%
  dplyr::select(village_number, localidade, lng, lat)

# Try to get locations for each localidade for livestock
avg_locations <- locations %>%
  mutate()
  group_by(localidade) %>%
  summarise(lng = mean(lng),
            lat = mean(lat))


# # Define function for converting a meta_instance_name from livestock
# to a household id in the census dataset
numeric_convert <- function(x){
  # Minimize
  x <- strsplit(x, '-')
  # Make numeric
  x <- lapply(x, function(z){as.character(as.numeric(z))})
  x <- lapply(x, function(z){paste0(z[1], '-', z[2], '-', z[3], '-', z[4])})
  x <- unlist(x)
  return(x)
}

# Join acd and eldo perm info
acd <-
  left_join(acd %>% 
              mutate(permid = numeric_convert(PERM_ID)), 
            eldo_perm %>% 
              mutate(permid = numeric_convert(new_permid)),
            by = 'permid')

# It would be (a) a heat map of livestock (only pigs, goats and poultry) at the district and (b) a heat map of malaria incidence in the cohort children finally (c) both combined in a single index. We could even try to get mosquito densities but lets start with just malaria + livestock.

# USE ACD data for both human distribution and animal distribution
# The relevant variables are main_animal_<animal name>

# Keep only relevant vars

acd <- acd %>%
  dplyr::select(
    permid,
         gender,
         cluster,
         longitude,
         latitude,
    chicken = MAIN_ANIMAL_CHICK,
         animals = MAIN_ANIMALS,
         cattle = MAIN_ANIMAL_CATTLE,
         pigs = MAIN_ANIMAL_PIGS,
         horses = MAIN_ANIMAL_HORSE,
         sheel = MAIN_ANIMAL_SHEEP,
         goats = MAIN_ANIMAL_COATS,
         FAMILY_ID,
         DEMOGRAFIC_DOB,
         HOME_WORK,
         HOME_WORK_TIME,
         HAD_MALARIA,
         INDUSTRY,
         MAIN_MALARIA_CONTROL_UNDERNET_SLEEP) %>%
  filter(!is.na(longitude),
         !is.na(latitude))
```

## Overall distribution of humans (point map)

```{r}
ggplot() +
  geom_polygon(data = mop2_fortified,
               aes(x = long,
                   y = lat,
                   group = group)) +
  theme_databrew() +
  coord_map() +
  geom_point(data = locations,
             aes(x = lng,
                 y = lat),
             color = 'red',
             alpha = 0.6,
             size = 0.1) +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

## Overall distribution of humans ("heat map")

```{r}
library(raster)
library(rasterVis)

heat_map <- function(the_map = mop2,
                     the_densities = locations,
                     griddy = 0.1){
  
  # Define the densities
  df_spatial <- the_densities %>% mutate(x = lng, y = lat)
  coordinates(df_spatial) <- ~x+y
  proj4string(df_spatial) <- proj4string(the_map)
  # First we need to create a uniform grid
  # Create a gridded dataframe with values 
  # for the entire range (bbox) of magude
  df_grid <- expand.grid(lng = seq(bbox(the_map)[1,1],
                                   bbox(the_map)[1,2],
                                   by = griddy), # make this smaller for better quality
                         lat = seq(bbox(the_map)[2,1],
                                   bbox(the_map)[2,2],
                                   by = griddy), # make this smaller for better quality
                         pos_score = NA)
  df_grid$latitude <- df_grid$lat
  df_grid$longitude <- df_grid$lng
  coordinates(df_grid) <- ~longitude+latitude
  proj4string(df_grid) <- proj4string(the_map)
  # # Visualize the grid
  # plot(the_map)
  # points(df_grid$lng, df_grid$lat, pch = ".")
  
  
  # Loop through each point in our grid
  for (i in 1:nrow(df_grid)){
    message(i)
    # Get distance from this grid point to every point in df_spatial
    distances <- spDistsN1(pts = df_spatial,
                           pt = df_grid[i,],
                           longlat = TRUE)
    # Define which are acceptably close (define a radius)
    close_enough <- which(distances <= 50)
    # Get a positivity score ( 1 means yes and 0 means no)
    # positivity <- stats::weighted.mean(
    #   x = df_spatial$pcr_mujer[close_enough],
    #   w = (1 / distances[close_enough])^2,
    #                                    na.rm = TRUE)
    positivity <- mean(1 / (distances[close_enough]^(1/4)), na.rm = TRUE)
    # Assign irs to the dataframe
    df_grid$pos_score[i] <- positivity
  }
  
  # Convert df_grid to raster
  temp <- df_grid@data %>% arrange(lng, lat)
  r <- rasterFromXYZ(temp[, c('lng', 'lat', 'pos_score')])
  # plot(r)
  # plot(the_map, add = TRUE)
  
  # Make grid smaller and then reconvert to raster
  proj4string(df_grid) <- proj4string(the_map)
  x <- over(df_grid, polygons(the_map))
  df_grid_small <- df_grid[!is.na(x),]
  # plot(df_grid_small,
  #      col = adjustcolor('black', alpha.f = 0.2))
  temp <- df_grid_small@data %>% arrange(lng, lat)
  r <- rasterFromXYZ(temp[, c('lng', 'lat', 'pos_score')])
  plot(r, legend = FALSE)
  plot(the_map, add = TRUE)
  
}
heat_map(griddy = 0.2)
# heat_map(griddy = 0.01)

# points(df_spatial,
#        col = adjustcolor("black", alpha.f = 0.2),
#        cex = 0.2)

```

## Overall distribution of livestock (point map)

```{r}
livestock_long
ggplot() +
  geom_polygon(data = mop2_fortified,
               aes(x = long,
                   y = lat,
                   group = group)) +
  theme_databrew() +
  coord_map() +
  # geom_point(data = livestock,
  #            aes(x = lng,
  #                y = lat),
  #            color = 'red',
  #            alpha = 0.6,
  #            size = 0.1) +
  labs(x = '',
       y = '') +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

## Overall distribution of livestock ("heat map")

```{r}
livestock_long <- 
  acd %>% 
  dplyr::select(longitude,
                latitude, 
                animals)
out_list <- list()
counter <- 0
for(i in 1:nrow(livestock_long)){
  this_row <- livestock_long[i,]
  n_animals <- this_row$animals
  if(!is.na(n_animals)){
    if(n_animals > 0){
      out_row <- tibble(longitude = this_row$longitude,
                        latitude = this_row$latitude,
                        animals = 1:n_animals)
      counter <- counter + 1
      out_list[[counter]] <- out_row
    }
  }
}
livestock_long <- bind_rows(out_list) %>%
  mutate(lng = longitude,
         lat = latitude)
heat_map(the_map = mop2,
         the_densities = livestock_long,
         griddy = 0.1)

```

# Map of livestock

# Map of malaria incidence in cohort children

# Map of livestock and malaria incidence combined into single index

# Pending: Map of mosquito densities

# Technical details

This document was produced on `r Sys.Date()` on a `r Sys.info()['sysname']` machine (release `r Sys.info()['release']`. To reproduce, one should take the following steps:

1. Clone the repository at https://github.com/databrew/bohemia

2. Populate the `analyses/livestock_distribution/data` directory with the following files: `Distribuicao de gado em Mopeia (1).xlsx` (emailed to team members from Charfudin Sacoor on May 27 2019);  

3. Populate the `analyses/livestock_distribution/data/from_eldo` directory with the following files (emailed from Eldo to team on June 10 2019):

```
├── Census_2016.csv
├── COST_ACD_Core_13-05-2019.csv
├── COST_ACD_Follow_Up_Core_13-05-2019.csv
├── COST_ACD_Follow_Up_V062017_Core_10-05-2019.csv
├── COST_Censo2017_Core.1.4.2019.csv
├── COST_CROSS_2017_Core_Final_11.04.2019.csv
├── COST_CROSS_2018_Core_Final_11.10.2018.csv
├── COST_Cross_2019_Core_10-06-2019.csv
├── COST_PCD_2019_Core_15-04-2019.csv
├── COST_PCD_V062017_Core_17-01-2019.csv
├── COST_SprayStatus_by_Village_Id_11.04.2019.EE.csv
├── Livestock_Mopeia.csv
└── Pcd1_core.csv
```

4. Download the `COST_Permids.New&Old_EE.csv` file, sent by Eldo to team on June 11 2019, into `data/from_eldo`.

5. "Render" (using `rmarkdown`) the code in `analysis/livestock_distribution/README.Rmd`

Any questions or problems should be addressed to joe@databrew.cc