---
title: "Mopeia wrd maps"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.path = 'figures/')
```

```{r}
library(tidyverse)
library(bohemia)
library(sp)
library(ggrepel)
library(ggthemes)
library(databrew)
library(extrafont)
library(leaflet)
library(webshot)
# # install_github("wch/webshot")
library(htmltools)
library(rgeos)
# library(mapview)

# Get health facilities
hf <- bohemia::health_facilities

# Get polygon
shp <- shp_poly <- bohemia::mopeia_hamlets

shp@data$id <- 1:nrow(shp)
shp <- fortify(shp)
shp_poly@data$id <- as.character(1:nrow(shp_poly))
shp <- left_join(shp, shp_poly@data %>% dplyr::select(id, village))
# Keep only Mopeia
hf <- hf %>%
  filter(district == 'Mopeia')

# Get roads, water, etc.
roads <- roads_shp <- bohemia::mopeia_roads
water <- water_shp <- bohemia::mopeia_water
waterways <- waterways_shp <- bohemia::mopeia_waterways
roads <- fortify(roads)
water <- fortify(water)
waterways <- fortify(waterways)
water <- bind_rows(water, waterways)
```

```{r}
hf$long <- hf$lng
chunk_list <- list()
for(i in 1:nrow(shp_poly)){
  this_shp_poly <- shp_poly[i,]
  this_id <- this_shp_poly$village
  this_shp <- shp %>% filter(village == this_id)
  this_hf <- hf %>% filter(!is.na(lng))
  coordinates(this_hf) <- ~lng+lat
  proj4string(this_hf) <- proj4string(this_shp_poly)
  this_hf <- this_hf[!is.na(over(this_hf, polygons(this_shp_poly))),]
  if(nrow(this_hf) > 0){
    this_hf <- data.frame(this_hf) %>%
      mutate(long = lng)
  } 
  
  this_water <- water_shp[!is.na(over(water_shp, polygons(this_shp_poly))),]
  this_water <- fortify(this_water)
  # these_roads <- intersect(roads_shp, this_shp_poly)
  # these_roads <- rgeos::gIntersects(spgeom1 = roads_shp, polygons(this_shp_poly))
  these_roads <- roads_shp[!is.na(over(roads_shp, polygons(this_shp_poly))),]
  these_roads <- fortify(these_roads)
  
  if(nrow(this_water) == 0){
    coords <- coordinates(this_shp_poly)
    this_water <- data.frame(long = coords[1],
                             lat = coords[2],
                             group  = 1)
      
  }
  if(nrow(these_roads) == 0){
    coords <- coordinates(this_shp_poly)
    these_roads <- data.frame(long = coords[1],
                             lat = coords[2],
                             group  = 1)
  }
  if(nrow(this_hf) == 0){
    coords <- coordinates(this_shp_poly)
    this_hf <- data.frame(long = coords[1],
                          lng = coords[1],
                             lat = coords[2],
                             group  = 1)
  }
  
  # Assign
  assign(paste0('this_shp_poly', i),
         this_shp_poly)
  assign(paste0('this_hf', i),
         this_hf)
  
  
  # Make maps
  assign(paste0('ga', i), {
    ggplot() +
    geom_polygon(data = this_shp,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'white',
               fill = 'grey') +
  geom_point(data = this_hf,
             aes(x = lng,
                 y = lat),
             size = 0.6,
             alpha = 0.6) +
  coord_map() +
  databrew::theme_simple() +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = this_id,
       subtitle = paste0('(', nrow(this_hf), ' health facilities)'))
  })
  
  assign(paste0('gb', i), {
    ggplot() +
      geom_polygon(data = shp,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'white',
               fill = 'grey') +
    geom_polygon(data = this_shp,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'white',
               fill = 'darkred',
               alpha = 0.6) +
  # geom_point(data = hf,
  #            aes(x = lng,
  #                y = lat),
  #            size = 0.3,
  #            alpha = 0.6) +
  coord_map() +
  databrew::theme_simple() +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = paste0(this_id, ' (location in district)'))
  })
  
   assign(paste0('gc', i), {
    ggplot() +
    geom_polygon(data = this_shp,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'white',
               fill = 'grey') +
  geom_point(data = this_hf,
             aes(x = lng,
                 y = lat),
             size = 0.6,
             alpha = 0.6) +
       geom_path(data = this_water,
                 aes(x = long,
                     y = lat,
                     group = group),
                 color = 'blue') +
       geom_path(data = these_roads,
                 aes(x = long,
                     y = lat,
                     group = group),
                 color = 'brown') +
  coord_map() +
  databrew::theme_simple() +
  labs(x = 'Longitude',
       y = 'Latitude',
       title = paste0(this_id, ' (with connecting roads/rivers)'),
       subtitle = paste0('Blue = water; darkred = roads'))
  })
   
   this_chunk <- 
    paste0(
      '\n\n## ', 
      this_id,
      '\n\n',
      '```{r}\n',
      "leaflet() %>%
        addProviderTiles(providers$OpenStreetMap) %>% 
    addPolygons(data = shp_poly,
              # fillColor = NA,
              fillOpacity = 0,
              stroke = 0.8,
              weight = 0.8,
              color = 'black') %>%
        addPolygons(data = this_shp_poly", i, ",
              # fillColor = NA,
              fillOpacity = 0.5,
              fillColor = 'red',
              stroke = 0.2,
              color = 'black',
              weight = 1)\n"
      ,
    '```\n',
      '```{r}\n',
      'ga', i, '\ngb', i, '\ngc', i, '\n```\n\n```{r}\n',
      "leaflet() %>%
        addProviderTiles(providers$Esri.WorldImagery) %>% 
        addPolygons(data =  this_shp_poly", i, ",
              # fillColor = NA,
              fillOpacity = 0,
              stroke = 0.2,
              color = 'white') %>%
  addCircleMarkers(data = this_hf", i, ",
  lng = this_hf", i, "$long,
  lat = this_hf", i, "$lat,
                   fillOpacity = 1,
                   color = 'red',
                   radius = 1)\n"
      ,
  
      '```'
    )
   chunk_list[[i]] <- this_chunk
}
chunks <- do.call('c', chunk_list)
```


```{r}
# chunks <- paste0("\n\n```{r}\n```\n")
chunks <- paste0(chunks, collapse = '\n')

# Write our order / child-calls to a doc
file_connection <- file('children.Rmd')
writeLines(paste0('---\noutput: html_document\n---\n\n', 
                  chunks), 
           file_connection)
close(file_connection)
```


```{r child='children.Rmd', echo = FALSE, eval = TRUE}
```

```{r, results='hide', echo = FALSE}
file.remove('children.Rmd')
```








