---
title: "Minicensus report"
author: "www.databrew.cc"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               # echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 9.64,
               fig.height = 5.9,
               fig.path = 'figures/')
source('global.R')
```


## Household IDs' matching with hamlet codes

```{r, echo = TRUE}
pd <- data$non_repeats
truth_vector <- substr(pd$hh_id, 1, 3) == pd$hh_hamlet_code
incorrect <- which(!truth_vector)
incorrect_fids <- pd$wid[incorrect]
```

Of the `r nrow(pd)` entries, `r length(incorrect)` had household ID numbers did not match the hamlet code. These were for fieldworkers `r paste0(incorrect_fids, collapse = ' and ')`. Below is the info in question:

```{r}
pd[incorrect,] %>%
  dplyr::select(wid, hh_hamlet_code, hh_id, instanceID) %>% DT::datatable
```

## Worker IDs within accepted range (1-100)

```{r, echo = TRUE}
pd <- data$non_repeats
truth_vector <- pd$wid <= 100
incorrect_a <- which(!truth_vector)
incorrect_b <- which(is.na(truth_vector))
incorrect <- unique(c(incorrect_a, incorrect_b))
incorrect_fids <- pd$wid[incorrect]
```

Of the `r nrow(pd)` entries, `r length(incorrect)` had worker ID numbers outside of the acceptable range. These were for fieldworkers `r paste0(incorrect_fids, collapse = ' and ')`. Below is the info in question:

```{r}
pd[incorrect,] %>%
  dplyr::select(wid, instanceID) %>% DT::datatable
```

## Discrepancy in numbers of residents 

```{r, echo = TRUE}
pd <- data$non_repeats
out_list <- list()
for(i in 1:nrow(pd)){
  # Retrieve residency info
  this_pd <- pd[i,]
  this_id <- this_pd$instanceID
  this_wid <- this_pd$wid
  n_residents <- this_pd$hh_member_num_residents
  n_non_residents <- this_pd$hh_member_num_non_residents
  # Get number of repeats pertaining to the residents
  this_repeat <- data$repeats$repeat_household_members_enumeration %>%
    filter(instanceID == this_id)
  # Put results together
  out <- tibble(instanceID = this_id,
                wid = this_wid,
                n_residents,
                n_non_residents,
                n_repeats = nrow(this_repeat))
  out_list[[i]] <- out
}
pd <- bind_rows(out_list)
# Check for matching
pd$ok <- pd$n_residents + pd$n_non_residents == pd$n_repeats
```

Of the `r nrow(pd)` entries, `r length(which(!pd$ok))` had mismatches in the number of residents in the initial question and the number of repeat entries for each resident. Below is data pertaining to these mismatches:

Table 1

```{r}
pd %>%
  filter(!ok) %>%
  dplyr::select(-ok) %>%
  DT::datatable
```


# Summary statistics

## Number of residents

```{r, fig.height = 10, fig.width = 10}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, hh_member_num_residents) %>%
  tally

ggplot(data = pd,
       aes(x = hh_member_num_residents,
           y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~hh_hamlet_code) +
  labs(x = 'Number of residents',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

## Number of non-residents

```{r, fig.height = 10, fig.width = 10}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, hh_member_num_non_residents) %>%
  tally

ggplot(data = pd,
       aes(x = hh_member_num_non_residents,
           y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~hh_hamlet_code) +
  labs(x = 'Number of non-residents',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

## Livestock owned per household

The following charts are intentionally shown as distributions (rather than singular values) so as to identify any discrepancies.

### Pigs < 6 weeks

```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, animal = hh_n_pigs_less_than_6_weeks) %>%
  tally

ggplot(data = pd,
       aes(x = animal,
           y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~hh_hamlet_code) +
  labs(x = 'Number of animal',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

### Pigs > 6 weeks

```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, animal = hh_n_pigs_greater_than_6_weeks) %>%
  tally

ggplot(data = pd,
       aes(x = animal,
           y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~hh_hamlet_code) +
  labs(x = 'Number of animal',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

### Cattle > 1 year

```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, animal = hh_n_cows_greater_than_1_year) %>%
  tally

ggplot(data = pd,
       aes(x = animal,
           y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~hh_hamlet_code) +
  labs(x = 'Number of animal',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

### Cattle < 1 year


```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, animal = hh_n_cows_less_than_1_year) %>%
  tally

ggplot(data = pd,
       aes(x = animal,
           y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~hh_hamlet_code) +
  labs(x = 'Number of animal',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

## Mosquito nets per household

### Mosquito nets by hamlet

```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, n_nets_in_hh) %>%
  tally

ggplot(data = pd,
       aes(x = n_nets_in_hh,
           y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~hh_hamlet_code) +
  labs(x = 'Number of nets',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

### Mosquito nets (overall)

```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(n_nets_in_hh) %>%
  tally

ggplot(data = pd,
       aes(x = n_nets_in_hh,
           y = n)) +
  geom_bar(stat = 'identity') +
  labs(x = 'Number of nets',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

## Deaths

### Overall deaths

```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(how_many_deaths) %>%
  tally %>%
  mutate(how_many_deaths = ifelse(is.na(how_many_deaths), 0, how_many_deaths))

ggplot(data = pd,
       aes(x = how_many_deaths,
           y = n)) +
  geom_bar(stat = 'identity') +
  labs(x = 'Number of deaths',
       y = 'Observations')
```

```{r}
DT::datatable(pd)
```

### Deaths by hamlet

```{r}
pd <- data$non_repeats
pd <- pd %>%
  group_by(hh_hamlet_code, how_many_deaths) %>%
  tally %>%
  mutate(how_many_deaths = ifelse(is.na(how_many_deaths), 0, how_many_deaths))

ggplot(data = pd,
       aes(x = how_many_deaths,
           y = n)) +
  geom_bar(stat = 'identity') +
  labs(x = 'Number of deaths',
       y = 'Observations') +
  facet_wrap(~hh_hamlet_code)
```

```{r}
DT::datatable(pd)
```

## Place of death

```{r, echo = TRUE}
pd <- data$repeats$repeat_death_info
pd <- pd %>%
  group_by(death_location) %>%
  tally
DT::datatable(pd)
```

## Time to fill form

```{r}
library(lubridate)
pd <- data$non_repeats
pd$end_time <- lubridate::as_datetime(pd$end_time)
pd$start_time <- lubridate::as_datetime(pd$start_time)
pd$time <- pd$end_time - pd$start_time
avg <- mean(pd$time, na.rm = T)
```

The average time to fill out the form was `r round(avg)` seconds. The below chart shows the distribution.

```{r}
ggplot(data = pd,
       aes(x = time)) +
  geom_density(fill = 'darkorange') +
  labs(x = 'Minutes',y = 'Density') +
  geom_vline(xintercept = avg, lty = 2)
```

The below table shows the time per fieldworker:

```{r}
pd %>%
  arrange(desc(time)) %>%
  dplyr::select(wid, seconds = time, instanceID) %>%
  mutate(seconds = round(seconds, digits = 1)) %>%
  DT::datatable()
```

