---
title: "Cluster generation from (incomplete) minicensus"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: "hide"
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               # echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 9.64,
               fig.height = 5.9,
               fig.path = 'figures/')
options(scipen=999)
```

```{r}
## Load libraries
library(bohemia)
library(ggplot2)
library(lubridate)
library(dplyr)
library(ggplot2)
library(sp)
library(raster)
library(ggthemes)
library(sf)
library(RColorBrewer)
library(readr)
library(tidyr)
library(leaflet)
library(rgeos)
# options(scipen = '999')
theme_set(databrew::theme_simple())
```


```{r}
if('data.RData' %in% dir()){
  load('data.RData')
} else {
  pd_moz <- load_odk_data(the_country = 'Mozambique',
                    credentials_path = '../../credentials/credentials.yaml',
                    users_path = '../../credentials/users.yaml')
  pd_tza <- load_odk_data(the_country = 'Tanzania',
                      credentials_path = '../../credentials/credentials.yaml',
                      users_path = '../../credentials/users.yaml')
  is_local <- FALSE
  library(DBI)
  library(RPostgres)
  source('../../rpackage/bohemia/R/app_functions.R')
  save(pd_moz,
       pd_tza,
       file = 'data.RData')
}
minicensus_main <- bind_rows(
  pd_moz$minicensus_main,
  pd_tza$minicensus_main
)
minicensus_people <- bind_rows(
  pd_moz$minicensus_people,
  pd_tza$minicensus_people
)
na_to_zero <- function(x){ifelse(is.na(x), 0, x)}
gps <- bohemia::gps

df_adjust <- function(df){
  df %>%
    mutate(n_households = ifelse(df$iso == 'TZA', n_households * 1,
                                 ifelse(df$iso == 'MOZ', n_households * 0.55, 
                                        NA)))
}

# source('global.R')
source('try_clusters.R')

# Define the number of clusters required of each type
n_required <- 49

# Get age and household details
ages <- 
  bind_rows(
    pd_moz$minicensus_people %>% mutate(country = 'Mozambique'),
    pd_tza$minicensus_people %>% mutate(country = 'Tanzania')
  ) %>%
  mutate(days_old = Sys.Date() - dob) %>%
  mutate(years_old = days_old / 365.25) %>%
  mutate(is_child  = ifelse(country == 'Mozambique',
                            years_old >= 0 & years_old <= 5,
                            years_old >= 0 & years_old <= 15)) %>%
  group_by(country) %>%
  summarise(children = length(which(is_child)),
            people = n()) %>%
  ungroup %>%
  mutate(percent_children = round(children / people * 100, digits = 2))

hh <- bind_rows(
  pd_moz$minicensus_main %>% mutate(country = 'Mozambique'),
  pd_tza$minicensus_main %>% mutate(country = 'Tanzania')
) %>%
  group_by(country) %>%
  summarise(avg_size = mean(hh_size))
```

```{r}
# Create a df based on minicensus
left <- minicensus_people %>%
  left_join(minicensus_main %>% dplyr::select(instance_id,
                                              country = hh_country)) %>%
  mutate(years_old = (Sys.Date() - dob)/ 365.25) %>%
   mutate(is_child  = ifelse(country == 'Mozambique',
                            years_old >= 0 & years_old <= 5,
                            years_old >= 0 & years_old <= 15)) %>%
  group_by(instance_id) %>%
  summarise(n_members = n(),
            n_children = length(which(is_child)))
df <-
  left_join(left,
            minicensus_main %>% dplyr::select(instance_id,
                                              country = hh_country,
                                              code = hh_hamlet_code,
                                              n_people = hh_size,
                                              location = hh_geo_location)) 
locs <- extract_ll(minicensus_main$hh_geo_location)
df$lng <- locs$lng; df$lat <- locs$lat
df <- left_join(df, bohemia::locations %>% dplyr::select(code, clinical_trial))
```


# Clustering based on minicensus-collected data {.tabset}

## Question

The question: How many clusters can we form with the data _already_ collected via the minicensus?

## Methods

Clusters will be formed using full data from the minicensus on the spatial distribution of hamlets and certain "parameters" (buffer distance, minimum number of children per cluster, minimum number and type of animals, etc.). Since the minicensus is not finished, we cannot do this for all areas. However, we can do it for the areas done so far, so as to get a sense of how the observed spatial distribution of households (and demographic details) translate into clusters.

## Parameters and assumptions

We consider a cluster to be "complete" if it has the following characteristics:
- At least 30 children (25 +5 to account for potential lost to follow-up)  
- NO ANIMAL FLOOR  
- At least 2 kilometers between clusters of different treatment assignments (ie, a 1km "buffer" for each cluster, except in the case that the nearby cluster is of identical treatment assignment)  

We consider that the number of clusters is "enough" if 147 are reached (ie, sufficient for the random assignment of all clusters to 3 assignment groups of n=49).

We allow for clusters to be formed by more than one hamlet, but we do not allow for the splitting of hamlets. 

We consider that a "child" is some between 5 or younger in Mozambique and 15 or younger in Tanzania. 


The clustering algorithm:  
- Starts at the hamlet of the district headquarters  
- Assess whether the hamlet is a complete cluster, then  
  - If it is a complete cluster, it moves on to the next hamlet and starts a new cluster  
  - If it is not a complete cluster, it adds the nearest hamlet to the same cluster (and so on)   
- Uses _pre-assigned_ (but random) treatment groups, so as to allow for the assignation of buffers only when necessary  

## Tanzania

Note: this is excluding hamlets in other clinical trial


```{r, echo = TRUE}
out <- try_clusters_hh_level(the_country = 'Tanzania',
                         include_clinical = TRUE,
                         minimum_households = 0,
                         minimum_children = 30,
                         minimum_humans = 0,
                         minimum_animals = 35,
                         minimum_cattle = 0,
                         minimum_pigs = 0,
                         minimum_goats = 0,
                         km = 2,
                         max_km_from_hq = 1000,
                         start_at_hq = FALSE,
                         df = df)
```

#### Summary

```{r, results='asis'}
cat(paste0(out$summary_text))
```

```{r}
keep_only_n <- function(cluster_xdf, n = n_required){
  done <- cluster_xdf %>% filter(complete_cluster)
  done <- done %>% mutate(dummy = 1) %>% group_by(assignment_group) %>%
    mutate(cs = cumsum(dummy)) %>%
    ungroup %>%
    filter(cs <= n_required)
}
out_minimal <- keep_only_n(out$cluster_xdf)
```

If we keep only the minimal number required (`r n_required` clusters per arm) and do not cense those in buffer areas, this is estimated to reduce the number of participating households from `r gps %>% filter(iso == 'TZA', clinical_trial == 0) %>% summarise(x = sum(n_households)) %>% .$x` to `r sum(out_minimal$n_households)`.


```{r}
out$map
```

## Mozambique


```{r, echo = TRUE}
out <- try_clusters(the_country = 'Mozambique',
  include_clinical = TRUE,
  interpolate_animals = TRUE,
  interpolate_humans = TRUE,
  humans_per_household = hh$avg_size[hh$country == 'Mozambique'],
  p_children = ages$percent_children[ages$country == 'Mozambique'],
  minimum_households = 0,
  minimum_children = 30,
  minimum_humans = 0,
  minimum_animals = 0,
  minimum_cattle = 0,
  minimum_pigs = 0,
  minimum_goats = 0,
  km = 2,
  max_km_from_hq = 1000,
  start_at_hq = TRUE,
  df = gps %>% mutate(Country = ifelse(iso == 'TZA', 'Tanzania', 'Mozambique'))  %>% df_adjust)
```


#### Summary

```{r, results='asis'}
cat(paste0(out$summary_text))
```


```{r}
keep_only_n <- function(cluster_xdf, n = n_required){
  done <- cluster_xdf %>% filter(complete_cluster)
  done <- done %>% mutate(dummy = 1) %>% group_by(assignment_group) %>%
    mutate(cs = cumsum(dummy)) %>%
    ungroup %>%
    filter(cs <= n_required)
}
out_minimal <- keep_only_n(out$cluster_xdf)
out_maximal <- out$cluster_xdf
```

IMPORTANT: UNABLE TO FORM MINIMUM NUMBER OF CLUSTERS. Only `r out$cluster_xdf %>% filter(complete_cluster) %>% nrow()` clusters total formed.

If we do not cense those in buffer areas, this is estimated to reduce the number of participating households from `r sum(out_maximal$n_households)` to `r sum(out_minimal$n_households)`.

```{r}
out$map
```


## More resources

- Source code for the hamlet-level clustering algorithm: [here](https://github.com/databrew/bohemia/blob/master/rpackage/bohemia/inst/shiny/operations/try_clusters.R) 
- Interactive web app for modifying parameters and re-running cluster algorithm: [here](https://bohemia.team/operations) ("Clustering" tab)


