---
params:
  data: !r data.frame(f = 'acéntõ', a = 1, b = 2, x = 4, xx=5,xy=6,xyz=8, yy=9,yz=11,zz=12) 
  workers_on_separate_pages: True
output:
  pdf_document:
    keep_tex: yes
    includes:
      in_header: consent_verification_list_preamble.sty
tables: yes
classoption: landscape
---

```{r setup, include=FALSE}
# output: 
#   pdf_document:
#     latex_engine: xelatex
#      pandoc_args: [
#      "-V", "classoption=twocolumn"
#    ]
#     template: pdf/layout.tex
#     includes:
#       in_header: pdf/preamble.sty
library(knitr)# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = F,
               fig.height = 6)
# knit_engines$set(asis = function(options) {
#   if (options$echo && options$eval) knit_child(text = options$code)
# })
```


```{r}
library(dplyr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
# Specify paramaters
for (i in 1:length(params)){
  assign(names(params)[i],
         params[[i]],
         env = .GlobalEnv)
}
# dfx <- params$data
# load('/tmp/dfx.RData')
# save.image('/tmp/dfx.RData')
dfx <- data
```

\pagenumbering{arabic} 

```{r}
chunk_list <- list()
dfx$wid <- unlist(dfx[,1])
fwids <- sort(unique(dfx$wid))
name1 <- names(dfx)[1]

if(workers_on_separate_pages){
  for(i in 1:length(fwids)){
  this_chunk <- 
    paste0(
      name1, ' ', fwids[i], '\n\n\n```{r}\n',
      "kable(dfx  %>% dplyr::filter(wid == ", fwids[i], ") %>% dplyr::select(-wid), 'latex',  
      longtable = T)  %>%   
  kable_styling(latex_options = c('repeat_header')) %>%
    kable_styling(font_size = 8)  %>%
  column_spec(1:2, width = '1.2cm') %>%
  column_spec(3, width = '1.7cm') %>%
  column_spec(4, width = '2cm') %>%
  column_spec(5:ncol(dfx), width = '2cm')\n```\n\\newpage\n"
    )
  chunk_list[[i]] <- this_chunk
}
} else {
  chunk_list[[1]] <- paste0(
    "All fieldworkers\n\n\n```{r}\n",
    "kable(dfx %>% dplyr::select(-wid), 'latex',  
      longtable = T)  %>%   
  kable_styling(latex_options = c('repeat_header')) %>%
    kable_styling(font_size = 8)  %>%
  column_spec(1:2, width = '1.2cm') %>%
  column_spec(3, width = '1.7cm') %>%
  column_spec(4, width = '2cm') %>%
  column_spec(5:ncol(dfx), width = '2cm')\n```\n\\newpage\n"
  )
}


chunks <- unlist(chunk_list)
chunks <- paste0(chunks, collapse = '\n')

file_connection <- file('children.Rmd')
writeLines(paste0('---\noutput: pdf_document\n---\n\n', 
                  chunks), 
           file_connection)
close(file_connection)
```

```{r child='children.Rmd'}
```

```{r, results='hide'}
file.remove('children.Rmd')
```



End.

