---
params:
  xdata: !r data.frame(n_hh = 400, n_teams = 2, id_limit_lwr = 1, id_limit_upr = 400) # data on the hamlet
  loc_id: 'MKU'
  enumeration: TRUE
  use_previous: FALSE
  enumerations_data: NULL
  minicensus_main_data: NULL
  refusals_data: NULL
  include_name: FALSE
  li: FALSE
output:
  pdf_document:
    keep_tex: yes
    includes:
      in_header: visit_control_sheet_preamble.sty
tables: yes
classoption: landscape
---

```{r setup, include=FALSE}
# output: 
#   pdf_document:
#     latex_engine: xelatex
#      pandoc_args: [
#      "-V", "classoption=twocolumn"
#    ]
#     template: pdf/layout.tex
#     includes:
#       in_header: pdf/preamble.sty
library(knitr)# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = F,
               fig.height = 6)
# knit_engines$set(asis = function(options) {
#   if (options$echo && options$eval) knit_child(text = options$code)
# })
```


```{r}
library(dplyr)
library(bohemia)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
# Specify paramaters
for (i in 1:length(params)){
  assign(names(params)[i],
         params[[i]],
         env = .GlobalEnv)
}
save.image('/tmp/image.RData')
x <- bohemia::locations
x <- x %>% filter(code %in% loc_id)
if(nrow(x) > 0){
  loc_name <- paste0(x$Ward, ', ', x$Village, ', ', x$Hamlet)
  loc_name <- paste0(sort(loc_name), collapse = '; ')
} else {
  loc_name <- ' '
}
include_name <- FALSE
```

Visit control sheet for `r paste0(paste0(loc_id, collapse = ';'), ifelse(length(loc_id) == 1, paste0(' (', loc_name, ')'), ''))`

\pagenumbering{gobble}


```{r}
lc <- loc_id
n_hh <- as.numeric(xdata$n_hh)
if(is.na(n_hh)){
  n_hh <- 1000
}
n_teams <- as.numeric(xdata$n_teams)
id_limit_lwr <- as.numeric(xdata$id_limit_lwr)
id_limit_upr <- as.numeric(xdata$id_limit_upr)

# Get country
country <- 'Mozambique'
if(lc[1] %in% locations$code[locations$Country == 'Tanzania']){
  country <- 'Tanzania'
}
```


```{r}
id_vals <- 1:n_hh
id_vals <- id_vals[id_vals %in% id_limit_lwr:id_limit_upr]
n_hh <- length(id_vals)

team_numbers <- rep(1:n_teams, each = round(n_hh / n_teams))
while(length(team_numbers) < n_hh){
  team_numbers <- c(team_numbers, team_numbers[length(team_numbers)])
}

while(length(team_numbers) > n_hh){
  team_numbers <- team_numbers[1:n_hh]
}


if(country == 'Tanzania'){
  out_list <- list()
  for(lcx in 1:length(lc)){
    this_out <- tibble(`HHID` = paste0(lc[lcx], '-', bohemia::add_zero(id_vals, n = 3)),
                  team = team_numbers)
    out_list[[lcx]] <- this_out
  }
  out <-bind_rows(out_list)

  left <- locations %>% filter(code %in% lc) %>% dplyr::select(code, District, Ward, Village, Hamlet)
  out$code <- substr(out$`HHID`, 1, 3)
  out = left_join(out, left, by = 'code') %>% dplyr::select(-code)
} else {
  out_list <- list()
  for(lcx in 1:length(lc)){
    this_out <- tibble(`Código do agregado` = paste0(lc[lcx], '-', bohemia::add_zero(id_vals, n = 3)),
                       team = team_numbers) %>%
      arrange(`Código do agregado`)
    out_list[[lcx]] <- this_out
  }
  out <-bind_rows(out_list)
}
```

```{r}
chunk_list <- list()
if(!li){
  chunk_list <- 'You must log in'
} else {
if(country == 'Mozambique'){
  if(enumeration){

    for(i in 1:n_teams){
      chunk_list[[i]] <-
        paste0("Team ", i, "\n\n\nNome/Código do Inquiridor:\n\n\n```{r}\n",
               "kable(out %>% filter(team == ", i, ") %>% tibble %>% dplyr::select(-team) %>% dplyr::mutate(`Nome de chefe de agregado` = '", "', `Localização do Numero de Agregado` = '",   ' ', "', `Data de enumeração` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size = 9) %>% kable_styling(latex_options = c('repeat_header')) %>% column_spec(1:4,width = '4.5cm')   \n```\n\\newpage\n")
    }
  } else {
    if(use_previous){
      # NON ENUMERATION
      # We use the previously enumerated data then and have to remove absences / etc
      refusals <- refusals_data %>%
        mutate(reason_no_participate = ifelse(reason_no_participate %in% c('SEM COMENTARIO',
                                                                           'He didnt want to do it',
                                                                           'Dont know'),
                                              'refused',
                                              'not_present')) %>%
        group_by(hh_id, reason_no_participate) %>%
        tally
      # Define those who should not be visited again
      remove_these <- refusals %>% filter(reason_no_participate == 'refused' | length(which(reason_no_participate == 'not_present')) >= 3) %>%
        .$hh_id
      # Define those which have already been mini-censed (and can therefore be removed)
      remove_these_mc <- minicensus_main_data$hh_id
      remove_these <- unique(c(remove_these, remove_these_mc))
      # Define those with previous absences
      previous_absences <- refusals %>%
        group_by(agregado = hh_id) %>%
        filter(length(which(reason_no_participate == 'refused')) == 0) %>%
        filter(reason_no_participate == 'not_present') %>%
        summarise(n = n()) 
      this_df <- enumerations_data %>% dplyr::filter(hamlet_code == lc) %>% dplyr::select(
        agregado, village, ward, hamlet, hamlet_code, localizacao_agregado, todays_date, chefe_name, wid, hamlet_code) %>%
        filter(!is.na(agregado)) %>%
        dplyr::distinct(agregado, .keep_all = TRUE) %>%
        left_join(previous_absences) %>%   
        mutate(`Ausências anteriores` = ifelse(is.na(n), 0, n)) %>%
        dplyr::select(-n) %>%
        filter(!agregado %in% remove_these) %>%
        arrange(agregado)

      if(!is.null(df)){
        if(!include_name & nrow(this_df) > 0){
          this_df$chefe_name <- ' '
        }
      }
      
      # Add team numbers
      nr = nrow(this_df)
      if(nr > 0){
        team_vals <- sort(((1:nr) %% n_teams) + 1)
        this_df$team <- team_vals
        
        for(i in 1:n_teams){
          chunk_list[[i]] <-
            paste0("Team ", i, "\n\n\nNome/Código do Inquiridor:\n\n\n```{r}\n",
                   "kable(this_df %>% filter(team == ", i, ") %>% tibble %>% dplyr::select(-team) %>% dplyr::select(`Data de enumeração` = todays_date, `ID agregado` = agregado, `Posto administrativo e Localidade` = ward, `Povoado` = village, `Bairro` = hamlet, `ID Bairro` = hamlet_code, `Nome de chefe de agregado` = chefe_name, `Ausências anteriores`) %>% dplyr::mutate(`Data da visita` = ' ') %>%  dplyr::mutate(`O chefe de agregado ou Chefe de agregado sustituto assino o consentimento informado?` = 'Sim__Não__', `Foi realizado o formulario?` = 'Sim__Não__',  `Se Não foi visitado ou entrevistado, explique o porque?` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size = 7) %>% kable_styling(latex_options = c('repeat_header'))  %>% column_spec(1:5, width = '1.5cm') %>% column_spec(6, width = '0.7cm') %>% column_spec(7, width = '1.3cm') %>% column_spec(8, width = '1.2cm') %>% column_spec(9, width = '1.4cm') %>% column_spec(10:11, width = '1.8cm') %>% column_spec(12, width = '2cm') \n```\n\\newpage\n")
        }
        
        
      } else {
        chunk_list[[1]] <- paste0('No previously enumerated households for ', lc)
      }
      
    } else {
      for(i in 1:n_teams){
        chunk_list[[i]] <-
          paste0("Team ", i, "\n\n\nNome/Código do Inquiridor:\n\n\n```{r}\n",
                 "kable(out %>% filter(team == ", i, ") %>% tibble %>% dplyr::select(-team) %>% dplyr::mutate(`Nome de chefe de agregado` = '", 
                 # this_df$chief_name,
                 " ", # currently blocking names from being stored
                 "', `Localização do Numero de Agregado` = '",  ' ',  "', `Data de enumeração` = ' ", as.character(Sys.Date()), "', `O chefe de agregado assino o consentimento informado?` = ' ', `O chede de agregado foi entrevistado?` = ' ', `Data da visita` = ' ', `Se Não foi visitado ou entrevistado, explique o porque?` = ' ', `Observações/Comentários` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size = 7) %>% kable_styling(latex_options = c('repeat_header')) %>% column_spec(1:9,width = '1.5cm')   \n```\n\\newpage\n")
      }
    }
    
    
    # # Get the odk data
    # pd <- odk_data$data
    # pd <- pd$non_repeats
    # pd <- pd %>%
    #   dplyr::filter(hamlet_code == lc)
    
    
  }
  
} else {
  for(i in 1:n_teams){
    chunk_list[[i]] <-
      paste0("Team ", i, "\n\n\nFieldworker name/code:\n\n\n```{r}\n",
             "kable(out %>% filter(team == ", i, ") %>% tibble %>% dplyr::select(-team) %>% dplyr::mutate(`Status` = ' ', `Comments` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size =11) %>% kable_styling(latex_options = c( 'repeat_header')) %>% column_spec(1:6, width = '2.8cm') %>% column_spec(7, width = '4cm')   \n```\n\\newpage\n")
  }
}
}


chunks <- unlist(chunk_list)
chunks <- paste0(chunks, collapse = '\n')

file_connection <- file('children.Rmd')
writeLines(paste0('---\noutput: pdf_document\n---\n\n', 
                  chunks), 
           file_connection)
close(file_connection)
```

```{r child='children.Rmd'}
```

```{r, results='hide'}
file.remove('children.Rmd')
```


End.