---
params:
  xdata: !r data.frame(n_hh = 400, n_teams = 2, id_limit_lwr = 1, id_limit_upr = 400) # data on the hamlet
  loc_id: 'MKU'
  enumeration: TRUE
  use_previous: FALSE
  enumerations_data: NULL
  include_name: FALSE
output:
  pdf_document:
    keep_tex: yes
    includes:
      in_header: visit_control_sheet_preamble.sty
tables: yes
classoption: landscape
---

```{r setup, include=FALSE}
# output: 
#   pdf_document:
#     latex_engine: xelatex
#      pandoc_args: [
#      "-V", "classoption=twocolumn"
#    ]
#     template: pdf/layout.tex
#     includes:
#       in_header: pdf/preamble.sty
library(knitr)# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = F,
               fig.height = 6)
# knit_engines$set(asis = function(options) {
#   if (options$echo && options$eval) knit_child(text = options$code)
# })
```


```{r}
library(dplyr)
library(bohemia)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
# Specify paramaters
for (i in 1:length(params)){
  assign(names(params)[i],
         params[[i]],
         env = .GlobalEnv)
}
# save.image('/tmp/image.RData')
x <- bohemia::locations
x <- x %>% filter(code == loc_id)
if(nrow(x) > 0){
  loc_name <- paste0(x$Ward, ', ', x$Village, ', ', x$Hamlet)
} else {
  loc_name <- ' '
}
```

Visit control sheet for `r paste0(loc_id, ' (', loc_name, ')')`

\pagenumbering{gobble}


```{r}
lc <- loc_id
n_hh <- as.numeric(xdata$n_hh)
n_teams <- as.numeric(xdata$n_teams)
id_limit_lwr <- as.numeric(xdata$id_limit_lwr)
id_limit_upr <- as.numeric(xdata$id_limit_upr)

# Get country
country <- 'Mozambique'
if(lc %in% locations$code[locations$Country == 'Tanzania']){
  country <- 'Tanzania'
}
```


```{r}
id_vals <- 1:n_hh
id_vals <- id_vals[id_vals %in% id_limit_lwr:id_limit_upr]
n_hh <- length(id_vals)

team_numbers <- rep(1:n_teams, each = round(n_hh / n_teams))
while(length(team_numbers) < n_hh){
  team_numbers <- c(team_numbers, team_numbers[length(team_numbers)])
}

while(length(team_numbers) > n_hh){
  team_numbers <- team_numbers[1:n_hh]
}


if(country == 'Tanzania'){
  out <- tibble(`HHID` = paste0(lc, '-', bohemia::add_zero(id_vals, n = 3)),
              team = team_numbers) 
  
  left <- locations %>% filter(code == lc) %>% dplyr::select(District, Ward, Village, Hamlet)
  out = bind_cols(left, out)
  } else {
  out <- tibble(`Código do agregado` = paste0(lc, '-', bohemia::add_zero(id_vals, n = 3)),
              team = team_numbers)
}
```

```{r}
chunk_list <- list()
if(country == 'Mozambique'){
  if(enumeration){
      this_df <- gps %>% dplyr::filter(code == lc) %>% left_join(locations %>% dplyr::select(code,Village, Hamlet))
  
  for(i in 1:n_teams){
  chunk_list[[i]] <-
    paste0("Team ", i, "\n\n\nNome/Código do Inquiridor:\n\n\n```{r}\n",
           "kable(out %>% filter(team == ", i, ") %>% tibble %>% dplyr::select(-team) %>% dplyr::mutate(`Nome de chefe de agregado` = '", "', `Localização do Numero de Agregado` = '",   ' ', "', `Data de enumeração` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size = 9) %>% kable_styling(latex_options = c('repeat_header')) %>% column_spec(1:4,width = '2.5cm')   \n```\n\\newpage\n")
}
  } else {
    if(use_previous){
      # NON ENUMERATION
    # We use the previously enumerated data then
      this_df <- enumerations_data %>% dplyr::filter(hamlet_code == lc) %>% dplyr::select(
        agregado, hamlet, hamlet_code, localizacao_agregado, todays_date, chefe_name, wid, hamlet_code) %>%
        filter(!is.na(agregado)) %>%
        dplyr::distinct(agregado, .keep_all = TRUE)
      if(!include_name & nrow(df) > 0){
        this_df$chefe_name <- ' '
      }
      # Add team numbers
      nr = nrow(this_df)
      if(nr > 0){
        team_vals <- sort(((1:nr) %% n_teams) + 1)
        this_df$team <- team_vals
        
        for(i in 1:n_teams){
            chunk_list[[i]] <-
          paste0("Team ", i, "\n\n\nNome/Código do Inquiridor:\n\n\n```{r}\n",
                 "kable(this_df %>% filter(team == i) %>% tibble %>% dplyr::select(-team) %>% dplyr::select(ID = agregado, `Localização do Numero de Agregado` = localizacao_agregado, `Data de enumeração` = todays_date, `Nome de chefe de agregado` = chefe_name) %>% dplyr::mutate(`O chefe de agregado assino o consentimento informado?` = ' ', `O chefe foi entrevistado?` = ' ', `Data da visita` = ' ', `Se Não foi visitado ou entrevistado, explique o porque?` = ' ', `Observações i Comentários` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size = 7) %>% kable_styling(latex_options = c('repeat_header')) %>% column_spec(1:3,width = '1.4cm') %>% column_spec(4,width = '3cm') %>% column_spec(5:7,width = '1.8cm') %>% column_spec(8:9,width = '2.4cm')   \n```\n\\newpage\n")
          }
        
        
      } else {
        chunk_list[[1]] <- paste0('No previously enumerated households for ', lc)
      }
      
    } else {
      for(i in 1:n_teams){
  chunk_list[[i]] <-
    paste0("Team ", i, "\n\n\nNome/Código do Inquiridor:\n\n\n```{r}\n",
           "kable(out %>% filter(team == ", i, ") %>% tibble %>% dplyr::select(-team) %>% dplyr::mutate(`Nome de chefe de agregado` = '", 
           # this_df$chief_name,
           " ", # currently blocking names from being stored
           "', `Localização do Numero de Agregado` = '",  ' ',  "', `Data de enumeração` = ' ", as.character(Sys.Date()), "', `O chefe de agregado assino o consentimento informado?` = ' ', `O chede de agregado foi entrevistado?` = ' ', `Data da visita` = ' ', `Se Não foi visitado ou entrevistado, explique o porque?` = ' ', `Observações/Comentários` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size = 7) %>% kable_styling(latex_options = c('repeat_header')) %>% column_spec(1:9,width = '1.5cm')   \n```\n\\newpage\n")
}
    }
    
  
    # # Get the odk data
    # pd <- odk_data$data
    # pd <- pd$non_repeats
    # pd <- pd %>%
    #   dplyr::filter(hamlet_code == lc)
    
  
  }

} else {
  for(i in 1:n_teams){
  chunk_list[[i]] <-
    paste0("Team ", i, "\n\n\nFieldworker name/code:\n\n\n```{r}\n",
           "kable(out %>% filter(team == ", i, ") %>% tibble %>% dplyr::select(-team) %>% dplyr::mutate(`Status` = ' ', `Comments` = ' '), 'latex', longtable = T, booktabs=FALSE, caption = 'Team ", i,"')  %>%   kable_styling(font_size =11) %>% kable_styling(latex_options = c( 'repeat_header')) %>% column_spec(1:6, width = '2.8cm') %>% column_spec(7, width = '4cm')   \n```\n\\newpage\n")
}
}

chunks <- unlist(chunk_list)
chunks <- paste0(chunks, collapse = '\n')

file_connection <- file('children.Rmd')
writeLines(paste0('---\noutput: pdf_document\n---\n\n', 
                  chunks), 
           file_connection)
close(file_connection)
```

```{r child='children.Rmd'}
```

```{r, results='hide'}
file.remove('children.Rmd')
```


End.