---
title: "Rufiji health post locations"
output: github_document
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8.64,
               fig.height = 4.86,
               fig.path = 'figures/')
```


**What**: What follows are some code examples for generating maps of Rufiji health facility locations. 

**Why**: Issa requested technical assistance from Joe to generate maps of Rufiji health posts. To the extent that both (a) the maps themselve and (b) the code for generating the maps may be useful to others involved in the Bohemia project, this document is being made public.

```{r}
# Libraries
library(Hmisc)
library(tidyverse)
library(readxl)
library(ggplot2)
library(haven)
library(bohemia)
library(sp)
library(ggthemes)
library(ggrepel)
# Read in rufiji data
hp <- read_excel('Health facilities in Rufiji Valley.xlsx')
names(hp) <- Hmisc::capitalize(names(hp))
# Classify the types of posts
hp <- hp %>%
  dplyr::mutate(type = ifelse(grepl('Hospital', `Facility name`),
                              'Hospital',
                              ifelse(grepl('Health Center', `Facility name`),
                                     'Health Center',
                                     ifelse(grepl('Dispensary', `Facility name`),
                                            'Dispensary', 'Other'))))
```

## Simple plot

```{r}
plot(bohemia::ruf2)
points(hp$Longitude, hp$Latitude)
```

## Slightly more beautiful plot

```{r}
rufex <- fortify(bohemia::ruf2, id = NAME_2)
ggplot() +
  geom_polygon(data = rufex,
       aes(x = long,
           y = lat,
           group = group),
       fill = 'brown',
       alpha = 0.2) +
  theme_map() +
    coord_map() +
  geom_point(data = hp,
             aes(x = Longitude,
                 y = Latitude))
```


## Map with labels

```{r}
# "Fortify" the map data
rufex <- fortify(bohemia::ruf2, id = NAME_2)
# Modify the point data
pd <- hp %>%
               dplyr::mutate(`Facility name` = gsub(' Hospital| Health Center| Dispensary', '', `Facility name`))
ggplot() +
  geom_polygon(data = rufex,
       aes(x = long,
           y = lat,
           group = group),
       fill = 'brown',
       alpha = 0.2) +
  theme_map() +
    coord_map() +
  geom_point(data = pd,
             aes(x = Longitude,
                 y = Latitude,
                 color = type)) +
  geom_text_repel(data = pd,
                  aes(x = Longitude,
                      y = Latitude,
                      label = `Facility name`),
                  size = 2) +
  scale_color_manual(name = '',
                     values = c('black','blue', 'red'))
```


## Map with labels and roads

```{r}
# "Fortify" the map data
rufex <- fortify(bohemia::ruf2, id = NAME_2)
roads <- fortify(bohemia::rufiji_roads, id = osm_id)
# Modify the point data
pd <- hp %>%
               dplyr::mutate(`Facility name` = gsub(' Hospital| Health Center| Dispensary', '', `Facility name`))
ggplot() +
  geom_polygon(data = rufex,
       aes(x = long,
           y = lat,
           group = group),
       fill = 'brown',
       alpha = 0.2) +
  theme_map() +
    coord_map() +
  geom_point(data = pd,
             aes(x = Longitude,
                 y = Latitude,
                 color = type)) +
  geom_text_repel(data = pd,
                  aes(x = Longitude,
                      y = Latitude,
                      label = `Facility name`),
                  size = 2) +
  geom_path(data = roads,
                   aes(x = long,
                       y = lat,
                       group = group),
            alpha = 0.7,
            color = 'black') +
  scale_color_manual(name = '',
                     values = c('black','blue', 'red')) 
```


## Map with labels, roads, and water

```{r}
# "Fortify" the map data
rufex <- fortify(bohemia::ruf2, id = NAME_2)
roads <- fortify(bohemia::rufiji_roads, id = osm_id)
water <- fortify(bohemia::rufiji_water, id = osm_id)
waterways <- fortify(bohemia::rufiji_waterways, id = osm_id)
# Modify the point data
pd <- hp %>%
               dplyr::mutate(`Facility name` = gsub(' Hospital| Health Center| Dispensary', '', `Facility name`))
ggplot() +
  geom_polygon(data = rufex,
       aes(x = long,
           y = lat,
           group = group),
       fill = 'brown',
       alpha = 0.2) +
  theme_map() +
  coord_map() +
  geom_point(data = pd,
             aes(x = Longitude,
                 y = Latitude,
                 color = type)) +
  geom_path(data = roads,
                   aes(x = long,
                       y = lat,
                       group = group),
            alpha = 0.7,
            color = 'black') +
  geom_polygon(data = water,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = 'blue',
               alpha = 0.7) +
    geom_path(data = waterways,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'blue',
               alpha = 0.7) +
  geom_text_repel(data = pd,
                  aes(x = Longitude,
                      y = Latitude,
                      label = `Facility name`),
                  size = 2) +
  scale_color_manual(name = '',
                     values = c('black','green', 'red')) 
```

## Interactive plot

The below generates an interactive plot. It's available for viewing at [http://databrew.cc/rufiji.html](http://databrew.cc/rufiji.html).

```{r}
library(leaflet)
library(leaflet.extras)
library(htmlwidgets)

l <- leaflet() %>%
  addProviderTiles(provider = providers$Esri.WorldImagery) %>%
  addPolygons(data = bohemia::ruf2,
              stroke = NA,
              color = 'red') %>%
  addMarkers(data = hp, lng = ~Longitude, lat = ~Latitude,
             popup = ~`Facility name`)
# l
htmlwidgets::saveWidget(l, file = 'rufiji.html', selfcontained = TRUE)
```

```{r, echo = FALSE}
# Run below from linux command line to host map
# cp rufiji.html ../../../databrew.github.io/
```

## Technical details


This document was produced on `r Sys.Date()` on a `r Sys.info()['sysname']` machine (release `r Sys.info()['release']`. To reproduce, one should take the following steps:

- Clone the repository at https://github.com/databrew/bohemia

- "Render" (using `rmarkdown`) the code in `analysis/rufiji_health_posts/README.Rmd`

Any questions or problems should be addressed to joe@databrew.cc
