---
title: "Bohemia insights, IHI Training"
author: "Databrew LLC"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8.64,
               fig.height = 4.86,
               fig.path = 'figures/')
```


## Context

In the context of training of census fieldworkers at IHI, (practice) data entry with the census form has begun. Since fieldworkers have already entered and uploaded data, Imani (Data Manager, IHI) requested the following from Databrew on Friday, March 13:

1. Average start and end time for the form

2. Tables and graphs of some useful data

3. Maps of the visited households and grazing areas and water bodies

This report addresses these requests. It includes both the outputs as well as the input code, for purposes of training and reproducibility.

## Data retrieval

```{r}
# In order for the below to function, you'll need to have correctly populated
# a credentials.yaml file, as per directions sent by Joe Brew in early March 2020
# Load libraries
library(bohemia)
library(dplyr)
library(leaflet)
library(ggplot2)
library(lubridate)
library(yaml)

# Get credentials
creds <- yaml.load_file('../../credentials/credentials.yaml')
url = creds$tza_odk_server
user = creds$tza_odk_user
pass = creds$tza_odk_pass

# Get form list
fl <- odk_list_forms(url = url,
                     user = user, 
                     pass = pass)
# Display
fl

# Extract the id of the form
id <- fl$id[fl$name == 'Census']

# Get all data for form in question (cache so as to not have to overping the server)
file_name <- 'cache.RData'
if(!file_name %in% dir()){
  
    # Get a list of the submissions for the form in question
  submissions <- odk_list_submissions(url = url,
                                      id = id,
                                      user = user, 
                                      password = pass)
  # View
  submissions

  # Get all the submissions
  data_list = odk_get_data(url = url, id = id, user = user, pass = pass)

  save(data_list, file = file_name)
} else {
  load(file_name)
}

# There is now an object called data_list, with a slot for repeats and non repeats

```

## 1. Time for form completion

```{r}
# Process the data
# We'll take the difference between start time and end time to get average
pd <- data_list$non_repeats
pd$start_time <- lubridate::as_datetime(pd$start_time)
pd$end_time <- lubridate::as_datetime(pd$end_time)
pd$time_diff <- as.numeric(difftime(pd$end_time, pd$start_time, units = 'mins'))
```


```{r}
# Average start and end time
mean(pd$start_time)

mean(pd$end_time)

paste0(mean(pd$time_diff), ' minutes')
```


```{r}
# Visualize the data

# Distribution of time to finish
ggplot(data = pd,
       aes(x = time_diff)) +
  geom_density() +
  databrew::theme_simple() +
  # add a line at the average time
  geom_vline(xintercept = mean(pd$time_diff), lty = 2) +
  labs(title = 'Time to complete census form')
```


```{r}
# Distribution of start time
ggplot(data = pd,
       aes(x = start_time)) +
  geom_histogram(fill = 'beige', color = 'grey') +
  databrew::theme_simple() +
  # add a line at the average time
  geom_vline(xintercept = mean(pd$start_time), lty = 2) +
  labs(title = 'Start time',
       labs = 'Time',
       y = 'Count')
```

```{r}
# Distribution of end time
ggplot(data = pd,
       aes(x = end_time)) +
  geom_histogram(fill = 'beige', color = 'grey') +
  databrew::theme_simple() +
  # add a line at the average time
  geom_vline(xintercept = mean(pd$end_time), lty = 2) +
  labs(title = 'End time',
       labs = 'Time',
       y = 'Count')
```

```{r}
# Copmbined distribution
x = pd 
x <- x %>% arrange(start_time)
x$wid <- factor(x$wid, levels = x$wid)
ggplot(data = x,
       aes(x = (wid),
           y = end_time)) +
  geom_segment(aes(yend = start_time,
           xend = (wid))) +
  geom_point() +
  geom_point(aes(y = start_time)) +
  databrew::theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8)) +
  labs(y = 'Time of day',
       x = 'Worker ID',
       title = 'Time to finish form by worker ID')
```

```{r}
# Time to finish by fieldworker
x = pd %>%
  group_by(wid = as.character(wid)) %>%
  summarise(n_forms = n(),
            avg_time = mean(time_diff))
x <- x %>% arrange(desc(avg_time))
x$wid <- factor(x$wid, levels = x$wid)
ggplot(data = x,
       aes(x = wid,
           y = avg_time)) +
  geom_bar(stat = 'identity', fill = NA, color = 'black') +
  databrew::theme_simple() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8)) +
  labs(y = 'Minutes',
       x = 'Worker ID',
       title = 'Time to finish form by worker ID')

```



## 2. Tables and graphs of some useful data

(Skipping for now due to time constraints)

## 3. Maps of the visited households and grazing areas and water bodies

#### Visited households

```{r}
# Prepare data
pd <- data_list$non_repeats

# Define function for extracting lng/lat
extract_geo <- function(x, give_back = c('lng', 'lat')){
  xs <- lapply(strsplit(x, ' '), function(z){tibble(lng = z[2], lat = z[1])})
  xs <- bind_rows(xs)
  xs$lng <- as.numeric(xs$lng)
  xs$lat <- as.numeric(xs$lat)
  if(length(give_back) == 2){
    return(xs)
  } else {
  return(unlist(xs[,give_back]))
  }
  
}
pd$lng <- extract_geo(pd$hh_geo_location, 'lng')
pd$lat <- extract_geo(pd$hh_geo_location, 'lat')
```

```{r}
# Simple map
library(sp)
plot(bohemia::ruf3)
points(pd$lng, pd$lat, col = 'red')
```

```{r}
# Zoomed in map
xy <- pd[,c('lng', 'lat')]
xy <- xy[!is.na(xy$lng),]
coordinates(xy) <- ~lng+lat
proj4string(xy) <- proj4string(ruf3)
keep <- sort(unique(sp::over(xy, polygons(bohemia::ruf3))))
sub_area <- ruf3[keep,]
plot(sub_area)
points(pd$lng, pd$lat, col = 'red')
```

```{r}
# Zoomed in with added details
plot(sub_area)
plot(rufiji_water, col = 'blue', add = T)
plot(rufiji_waterways, col = 'blue', add = T)
plot(rufiji_roads, col = 'black', add = T)
points(pd$lng, pd$lat, col = 'red', pch = '.')
```

```{r}
# Interactive
leaflet() %>%
  addProviderTiles('Esri.WorldImagery') %>%
  addMarkers(lng = pd$lng,
             lat = pd$lat,
             popup = paste0('Worker ID: ', pd$wid, ' | Form ID: ', pd$instanceID))
```

#### Grazing areas

(Nothing geocoded)



```{r}
gr <- data_list$repeats$repeat_water
gr$lng <- extract_geo(gr$water_bodies_geo, 'lng')
gr$lat <- extract_geo(gr$water_bodies_geo, 'lat')

leaflet() %>%
  addProviderTiles('Esri.WorldImagery') %>%
  addCircleMarkers(lng = gr$lng,
             lat = gr$lat,
             popup = paste0(gr$water_bodies_type, ' | Form ID: ', pd$instanceID))
```


#### Bodies of water
