---
title: "Household head consent issues"
subtitle: "Mozambique"
author: "www.databrew.cc"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html:
    toc: true
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


```{r include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8.64,
               fig.height = 4.86,
               fig.path = 'figures/')
```


```{r}
# Libraries
library(Hmisc)
library(tidyverse)
library(readxl)
library(ggplot2)
library(haven)
library(bohemia)
library(sp)
library(ggthemes)
library(ggrepel)
```



```{r}
if('data.RData' %in% dir()){
  load('data.RData')
} else {
  data <- load_odk_data(the_country = 'Mozambique',
                        credentials_path = '../../credentials/credentials.yaml',
                        users_path = '../../credentials/users.yaml')
  save(data, file = 'data.RData')
}
```


```{r}
pd <- data$minicensus_main %>%
  mutate(hh_head_id = as.numeric(hh_head_id)) %>%
  mutate(respondent_id = as.numeric(respondent_id)) %>%
  dplyr::select(instance_id, hh_head_id,
                n_subs = hh_sub_size,
                n_residents = hh_member_num_residents,
                respondent_id,
                n_non_residents = hh_member_num_non_residents,
                todays_date) %>%
  mutate(head_responds = respondent_id == hh_head_id) 
# Get info on the subs
subs_info <- data$minicensus_repeat_hh_sub %>%
  left_join(pd %>% dplyr::select(instance_id, respondent_id)) %>%
  group_by(instance_id) %>%
  summarise(sub_responds = any(respondent_id == hh_sub_id))
# Join information on heads so subs
joined <- left_join(pd, subs_info) %>%
  mutate(sub_responds = ifelse(is.na(sub_responds), FALSE, sub_responds)) %>%
  mutate(neither_sub_nor_head = !sub_responds & !head_responds) 
# Get info on household heads
heads <- data$minicensus_people %>%
  dplyr::select(instance_id, first_name, last_name, pid,
                num, dob, gender) %>%
  left_join(pd %>% dplyr::select(instance_id,
                                 hh_head_id)) %>%
  filter(hh_head_id == num) %>%
  mutate(initials = paste0(first_name, last_name))
# Join other info
out <- left_join(heads,
                   joined %>% dplyr::select(date = todays_date,
                                            instance_id,
                                            n_subs,
                                            n_residents,
                                            n_non_residents,
                                            head_responds,
                                            sub_responds,
                                            neither_sub_nor_head)) %>%
  mutate(status = ifelse(head_responds, 'Head answered questions',
                   ifelse(sub_responds, 'Sub answered questions',
                          'Neither'))) %>%
    left_join(data$minicensus_main %>% dplyr::select(instance_id, Hamlet = hh_hamlet, 
                                                     hh_id,
                                                   `Code` = hh_hamlet_code, 
                                                   `Village` = hh_village, 
                                                   `Ward` = hh_ward))
out <- out %>% dplyr::select(
  Ward, Village, Hamlet, Code,
  `Agregado` = hh_id, 
  Date = date,
  `Number of substitutes` = n_subs,
  `Residents` = n_residents,
  `Non-residents` = n_non_residents,
  Status = status,
  `Chefe birthday` = dob,
  `Chefe` = pid,
    `Chefe initials` = initials
) %>%
  mutate(`ICF signed by` = '______') %>%
  mutate(`Confirmed OK` = FALSE)

prettify(out,
         download_options = TRUE,
         nrows = nrow(out))
write_csv(out,
          'fix_table.csv')
```


# Technical details and caveats

- This report is as of `r Sys.time()`. The database is updated as new data streams in, so these numbers are subject to change.^[ie, a form may have been collected yesterday but not yet uploaded to the server.]  
- The data shown herein are as per the current status in the study database. There are pending corrections to be implemented (such as when a hamlet code is set to "XXX", or when a "practice" form was accidentally submitted as if it were real).  
- The code used for the generation of this report has _not_ been externally code-reviewed.^["Code review" refers to someone other than the author of the code reviewing line by line to ensure accuracy. Generally, it is a rule that a report's code is reviewed prior to sharing with a client; for the sake of speed, we have broken this rule.] Code review is expected finish by Monday October 19th.
- This document was produced on `r Sys.Date()` on a `r Sys.info()['sysname']` machine (release `r Sys.info()['release']`.  


Any questions or problems should be addressed to bohemia@databrew.cc
