---
title: "Coefficient of variation"
output: github_document
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 8.64,
               fig.height = 4.86,
               fig.path = 'figures/')
```


```{r}
# Libraries
library(tidyverse)
library(readxl)
library(ggplot2)
library(haven)
# Read in rufiji data
ruf <- read_excel('data/Malaria incidence rates -geographical variations from HDIS2 data.xlsx')
names(ruf) <- c('ward', 'population', 'cases', 'incidence')
# ruf2 <- read_excel('data/Malaria incidence rates- age groups.xlsx')
# Remove the total row
ruf <- ruf %>% filter(ward != 'OVERALL')

# Mopeia data
source('prepare_acd_data.R')
```

## The objective

We want a general notion of the the coefficient of variation (CV) in the incidence of malaria in the districts of Rufiji (Tanzania) and Mopeia (Mozambique).

## The challenges  

Estimating this with the currently available data is not straightforward. In addition to the obvious shortcomings in using data from several years ago (things may have changed since then), there is a larger limitation: the data we have from Mopeia is at the person-level, whereas the Rufiji data is aggregated at the level of the ward (the Tanzanian equivalent of the posto administrativo).

One could simply take the CV in the Mopeia clusters and compare it with the CV in the Rufiji wards, but this comparison would be inappropriate since the sizes of these areas are so radically different. Whereas Mopeia clusters are made up of several hundred people, Rufiji wards include thousand (and in some cases, tens of thousands).

A larger unit size inevitably leads to a smaller CV.



```{r}
ggplot(data = ruf,
       aes(x = population)) +
  geom_density(fill = 'darkorange', alpha = 0.3) +
  theme_databrew() +
  scale_y_continuous(breaks = seq(0,1, by = 0.00001)) +
  labs(x = 'Population of ward',
       y = 'Density')
```

## The work-around

Given the incompatabilities in the data being compared, we need to get creative. We previously estimated that CV for Mopeia, based on person-level data. We want to use similar methods on our two datasets, but their format (person-level vs. ward-level) is too dissimilar. Since we can't transform ward-level data into person-level data, we'll instead do the opposite: transorm our person-level data (from Mopeia) to pseudo-ward-level data (ie, much larger numbers than clusters), using resampling. 

The result will be a _true_ coefficient of variation based on observed Rufiji ward-level data and an _estimated_ coefficient of variation based on Mopeia pseudo-ward-level data (via resampling). To the extent that the two are similar, we can feel comfortable using the same CV for Rufiji as the one we estimated based on person-level data from Mopeia. To the extent that they are very different, we may need to consider getting more granular data from Rufiji before moving forward.

## Methods

```{r, echo = TRUE}
# Loop throgh each ward to get population size, resample from mopeia, calculate cv
acd$id <- 1:nrow(acd)
out <- rep(NA, nrow(ruf))
for(i in 1:nrow(ruf)){
  this_ward_size <- round(ruf$population[i])
  sampled <- acd[sample(acd$id, this_ward_size, replace = TRUE),]
  sampled_incidence <- length(which(sampled$HAD_MALARIA == 1)) / nrow(sampled)
  out[i] <- sampled_incidence
}

cv_rufiji <- sd(ruf$incidence) / mean(ruf$incidence)
cv_mopeia <- sd(out) / mean(out)
```


## Results

The ward-level coefficient of variation of the incidence of Malaria in Rufiji is: `r round(cv_rufiji, digits = 2)`.

The pseudo-ward-level coefficient of variation in the incidence of Malaria in Mopeia is: `r round(cv_mopeia, digits = 2)`.

## Interpretation

The results are not reliable. The number of observations in the Mopeia ACD dataset (approximately 1500) is far too small to be compared with the level of aggregation in the Rufiji incidence dataset (many thousand per ward). In order to "upscale" the Mopeia dataset, we (inevitably) have to resample from the small (1500) ACD numbers multiple times. Resampling from the same set many times artificially deflates variation, hence the radically different CV estimates between the two datasets.

## Comparability

Beyond the possible comparability of the two locations (which this exercise has failed to address), it is worth noting that in the Rufiji ward-level incidence data, variance is _extremely high_. The coefficient of variation in the ward-level data is `r cv_rufiji`, significantly higher than the CV we (Aina) previously estimated for Mopeia. Given this high ward-level CV, if we were to obtain more granular data (ie, geographical units at the same size of the Mopeia clusters) we should anticipate an _even higher_ CV. The take-away: variation is much higher in Rufiji than in Mopeia.

## Action

All of the prior information should be taken with a grain of salt. The data we have for Rufiji and Mopeia are _qualitatively_ different (passive vs. active case-finding, all ages vs. children, aggregated vs. granular) and therefore cannot be _quantitatively_ compared reliably. 

We should obtain granular (person-level) incidence data from Malaria, if it exists.


## Technical details


This document was produced on `r Sys.Date()` on a `r Sys.info()['sysname']` machine (release `r Sys.info()['release']`. To reproduce, one should take the following steps:

- Clone the repository at https://github.com/databrew/bohemia

- Populate the `analyses/cv/data` directory with the following files:

```
├── Malaria incidence rates- age groups.xlsx
└── Malaria incidence rates -geographical variations from HDIS2 data.xlsx
```

- Populate the `analyses/cv/data/from_eldo` directory with the following files (emailed from Eldo to team on June 10 2019):

```
├── Census_2016.csv
├── COST_ACD_Core_13-05-2019.csv
├── COST_ACD_Follow_Up_Core_13-05-2019.csv
├── COST_ACD_Follow_Up_V062017_Core_10-05-2019.csv
├── COST_Censo2017_Core.1.4.2019.csv
├── COST_CROSS_2017_Core_Final_11.04.2019.csv
├── COST_CROSS_2018_Core_Final_11.10.2018.csv
├── COST_Cross_2019_Core_10-06-2019.csv
├── COST_PCD_2019_Core_15-04-2019.csv
├── COST_PCD_V062017_Core_17-01-2019.csv
├── COST_SprayStatus_by_Village_Id_11.04.2019.EE.csv
└── Pcd1_core.csv
```

- Download the `COST_Permids.New&Old_EE.csv` file, sent by Eldo to team on June 11 2019, into `data/from_eldo`.

- "Render" (using `rmarkdown`) the code in `analysis/cv/README.Rmd`

Any questions or problems should be addressed to joe@databrew.cc
